---
title: "RQ 2 Linear Models List (Claude)"
format: docx
editor: visual
---

```{=html}
<!--
CORRECTED (this session): two discrepancies fixed against the actual
current Research_question_2_FIXED (Claude).qmd:
1. usgsthesis_model_month_glmm was missing "* Site" -- the real formula
   is Study.Dataset * Site + Month, matching the lmer version right
   above it. As typed here before, it only had Study.Dataset + Month
   (main effects, no interaction) -- a different, weaker model than the
   one actually in the file.
2. usgsthesis_model2_phenperiod_glmm (the PRIMARY model for that
   comparison) was missing entirely -- only its lmer sensitivity twin,
   usgsthesis_model2_phenperiod, was listed.
Everything else in this file was already correct -- Loc_ID was present
throughout, unlike the RQ1 and RQ3 tracking docs.
-->
```

#### Research Question 2: Does sediment deposition within the same marsh (ramped versus scarped edge marsh typology) change across years?

usgsthesis_model_month = lmer(Flux\~Study.Dataset\*Site + Month + (1\|Site:Tran) + (1\|Loc_ID))

usgsthesis_model_month_glmm = glmmTMB(Flux\~Study.Dataset \* Site + Month + (1\|Site:Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

usgsthesis_model_phenperiod = lmer(Flux \~ Study.Dataset + Phen_Period + Study.Dataset \* Phen_Period + Site + (1\|Site:Tran) + (1\|Loc_ID))

usgsthesis_model_phenperiod_glmm = glmmTMB(Flux \~ Study.Dataset + Phen_Period + Study.Dataset \* Phen_Period + Site + (1\|Site:Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# uses thesis_rq2

summary(usgsthesis_model_month)
summary(usgsthesis_model_month_glmm)
Anova(usgsthesis_model_month_glmm, type = "II")
emmeans(usgsthesis_model_month_glmm, pairwise ~ Study.Dataset | Site, type = "response")
summary(usgsthesis_model_phenperiod)
summary(usgsthesis_model_phenperiod_glmm)
```

usgsthesis_model2_month = lmer(Flux \~ Site + (1\|Study.Dataset) + (1\|Month)) *\# variance-partitioning check only -- Study.Dataset as a 2-level random effect can't reliably estimate a variance component; not the model to report for the USGS-vs-Thesis comparison*

usgsthesis_model2_month_glmm = glmmTMB(Flux\~ Site + (1\|Study.Dataset) + (1\|Month), family = Gamma(link = "log")) *\# same caveat as above*

usgsthesis_model2_phenperiod_glmm = glmmTMB(Flux \~ Site + Study.Dataset + (1\|Site:Tran) + (1\|Loc_ID), family = Gamma(link = "log")) *\# PRIMARY model for Model 2 -- Study.Dataset as fixed, since it's the actual comparison variable*

usgsthesis_model2_phenperiod = lmer(Flux \~ Site + Study.Dataset + (1\|Site:Tran) + (1\|Loc_ID)) *\# sensitivity check for usgsthesis_model2_phenperiod_glmm*

```{r}
# uses thesis_rq2
# NOTE: Study.Dataset only has 2 levels -- watch for a singular-fit 
# warning on usgsthesis_model2_month/_glmm

summary(usgsthesis_model2_month)
summary(usgsthesis_model2_month_glmm)
summary(usgsthesis_model2_phenperiod_glmm)
summary(usgsthesis_model2_phenperiod)
```

marsh_dist_model = lmer(log(Flux) \~ Study.Dataset \* Month \* Dist + (1\|Site/Tran) + (1\|Loc_ID))

marsh_dist_model_glmm = glmmTMB(Flux\~Study.Dataset \* Month \* Dist + (1\|Site/Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

marsh_area_dataset_model = lmer(log(Flux) \~ Marsh_Area \* Study.Dataset + Site + (1\|Site:Tran) + (1\|Loc_ID))

marsh_area_dataset_model_glmm = glmmTMB(Flux \~ Marsh_Area \* Study.Dataset + Site + (1\|Site:Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# uses thesis_rq2, filtered to Flux > 0
# marsh_dist_model(+_glmm) has no Site term -- (1|Site/Tran) nesting
# marsh_area_dataset_model(+_glmm) has Site as fixed -- (1|Site:Tran) nesting

summary(marsh_dist_model)
summary(marsh_dist_model_glmm)
summary(marsh_area_dataset_model)
summary(marsh_area_dataset_model_glmm)
```

bay_channel_model_month = lmer(Flux \~ Location + Month + Location \* Month + Site + (1\|Site:Tran) + (1\|Loc_ID))

bay_channel_model_month_glmm = glmmTMB(Flux\~Location + Month + Location \* Month + Site + (1\|Site:Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# uses bay_channel (built from usgs_bay_byrep.csv / usgs_river_byrep.csv)
# Site kept FIXED here (only 2 sites -- CorteMadera, SanPablo)

summary(bay_channel_model_month)
summary(bay_channel_model_month_glmm)
```

```{=html}
<!--
NOT YET IN THIS FILE, worth adding: Research_question_2_FIXED.qmd
currently has zero DHARMa diagnostics (no simulateResiduals() calls
anywhere) -- flagged earlier in the session, still unresolved. Worth
adding simulateResiduals()/plot()/testDispersion() for whichever of
these ends up primary before treating its p-values as final.
-->
```
