---
title: "RQ 3 Linear Models and Regression Lists (Claude)"
format: docx
editor: visual
---

```{=html}
<!--
CORRECTED (this session):
1. Every formula below was missing (1|Loc_ID) -- fixed to match the
   actual current RQ3-variables (Claude).qmd.
2. deposition_model_slope/base (+_glmm) are coded in RQ1-spatial.Rmd,
   not RQ3-variables.qmd -- kept tracked here on purpose, since the
   question they answer (does intra-marsh elevation/slope predict Flux)
   maps to RQ3's framing rather than RQ1's. Noted below which file
   actually runs them.
3. The original code chunk here had malformed/escaped backticks
   (```` \`\`\`{r} \# ... print(...)} ````) that would fail to run as
   written -- rewritten as a normal chunk below.
-->
```

## Research Question 3: Do intra-marsh characteristics have more of an effect on sediment deposition than marsh typology?

deposition_model_slope = lmer(Flux\~z_star + local_slope + (1\|Site) + (1\|Loc_ID))

deposition_model_base = lmer(Flux\~z_star + (1\|Site) + (1\|Loc_ID))

deposition_model_slope_glmm = glmmTMB(Flux\~z_star + local_slope + (1\|Site) + (1\|Loc_ID), family = Gamma(link= "log"))

deposition_model_base_glmm = glmmTMB(Flux\~z_star + (1\|Site) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# CODED IN RQ1-spatial.Rmd, tracked here by design -- see note above.
# uses master_tier5_byrep_clean
# base vs. slope compared via anova() -- nested model comparison, 
# does adding local_slope improve fit over z_star alone

summary(deposition_model_slope)
summary(deposition_model_base)
anova(deposition_model_base, deposition_model_slope)
summary(deposition_model_slope_glmm)
summary(deposition_model_base_glmm)
anova(deposition_model_base_glmm, deposition_model_slope_glmm)
```

rq3_model = lmer(log(Flux) \~ z_star + local_slope + marsh_edge_type + (1\|Site) + (1\|Loc_ID))

rq3_model_glmm = glmmTMB(Flux\~z_star + local_slope + marsh_edge_type + Dist + (1\|Site/Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# uses master_tier5_byrep_clean
# rq3_model is the Gaussian SENSITIVITY check; rq3_model_glmm is PRIMARY

summary(rq3_model)
summary(rq3_model_glmm)
Anova(rq3_model_glmm, type = "II")
```

rq3_model_glmm_edgeID = glmmTMB(Flux \~ z_star + local_slope + marsh_edge_type + Dist + (1\|marsh_edgeID) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# uses master_tier5_byrep_clean
# explicit COMPARISON against rq3_model_glmm, not a replacement --
# marsh_edgeID is interaction(Site, Tran) relabeled, not additional
# independent replication

summary(rq3_model_glmm_edgeID)
AIC(rq3_model_glmm, rq3_model_glmm_edgeID)
```

rq3_model_glmm_phenperiod = glmmTMB(Flux\~z_star + local_slope + marsh_edge_type + Dist + Phen_Period + (1\|Site/Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

rq3_model_glmm_season = glmmTMB(Flux\~z_star + local_slope + marsh_edge_type + Dist + Cal_Season + (1\|Site/Tran) + (1\|Loc_ID), family = Gamma(link = "log"))

```{r}
# uses master_tier5_byrep_clean
# each is compared against the base rq3_model_glmm separately --
# Phen_Period and Cal_Season are two groupings of the same underlying
# time variable, not combined into one formula together

summary(rq3_model_glmm_phenperiod)
summary(rq3_model_glmm_season)
AIC(rq3_model_glmm, rq3_model_glmm_phenperiod, rq3_model_glmm_season)
```
