---
title: "Independent Project Introduction"
author: "Savannah K. Miller"
date: "2025-03-09"
output: pdf_document
---

## Dataset Card Source: (https://github.com/Smillersfsu/DepositionSFB/blob/0a6b8b2033b76344b3b1574f88ab81aa244e5210/DatasetCard.md)

## Background

Sediment is the structural foundation of the salt marsh ecosystem where endemic species live and acquire resources(Levin et al. 2001; Janousek et al. 2017). The salt marsh ecosystem is host to many endemic species that are specially adapted to the conditions of the salt marsh(Dunson and Travis 1994; Schile et al. 2011; Moffett, Robinson, and Gorelick 2010; Rankin et al. 2023), and the very sediment they walk and live within is in short supply(Schoellhamer and Marineau, n.d.). Salt marshes play strong roles in dampening impacts of atmospheric rivers and extreme weather events, offering a protection to coastal communities(Smolders et al. 2015; Castagno et al. 2022; Lee and Nepf 2024; Taylor-Burns et al. 2024). Already, there has been a long human history of disturbance to existing salt marshes, which puts ecological and human communities at severe risk(Gedan, Silliman, and Bertness 2009). Under these circumstances, the few salt marshes left are at risk of elimination due to drowning if action is not taken and the benefits of this ecosystem could be forever lost.(Endris et al. 2024).

Across the state, salt marshes are decreasing in number and area due to a variety of factors(Endris et al. 2024). This unfortunate circumstance paired with changing water levels along the coastline increases the sense of urgency to understand, conserve, and restore California’s salt marshes. To assist in raising the elevation of the salt marshes to adapt to natural hazards, we need to understand the fundamental processes of how sediment accumulates within the salt marsh over a brief period (short-term deposition) (Fagherazzi 2013; Houttuijn Bloemendaal et al. 2021; Vandenbruwaene et al. 2011). In this study, I will measure short-term deposition rates across marsh edges of multiple marshes within a single estuary in San Pablo Bay of the San Francisco Bay Estuary and comparing those rates to measurements of the surrounding marsh landscape. This study will provide information to ongoing restoration efforts of salt marshes in San Pablo Bay.(Callaway 2024; University of San Francisco et al. 2011; K. M. Thorne et al. 2019; Haltiner et al. 1996) My study provides novel insights for land managers to characterize their marshes edges to adapt to natural hazards and enhance community resiliency(Reed et al. 1999).

In this study, I will measure short-term deposition rates on multiple marshes within the San Francisco Bay Estuary and will compare those rates to measurements of the surrounding marsh landscape. This study builds on the work of Karen Thorne and U.S. Geological Survey (2025; Inter-and Intra-Annual Sediment Dynamics in Two Tidal Marshes: Deposition, Accretion, and Vegetation Data - ScienceBase-Catalog) to understand differences in sediment deposition on marshes with different edge characteristics, vegetation densities and geographic settings. Moreover, I will measure variables such as elevation, vegetation structure, and observations of marsh edge traits and comparing to rates of short-term deposition measured to understand patterns between variables. Other environmental variables such as season at time of collection and tide level at time of collection will be recorded and considered in analyses.

## Research Objective

This study will measure short-term deposition rates of sediment across the exposed (bay-adjacent) salt marshes in the northern part of the San Francisco Bay Estuary of the San Pablo Bay. I will compare these rates to biological structures of each salt marsh.

### Research Aim:

Marsh characteristics of the vegetation also are one of the first interactions with the waves bay ward and along the marsh edge. These vegetation characteristics will be measured at each marsh and compared with the rates of sediment deposition measured over the course of a year’s spring-neap tidal cycles.

#### Expected Outcome:

Deposition rates vary significantly with characteristics of the marsh vegetation, specifically maximum height.

# Null Hypothesis:

Deposition rates of sediment do not vary with differences in height of Salicornia pacificia (Pickleweed).

# Alternate hypothesis

Deposition rates of sediment do vary with differences in height of Salicornia pacifica (Pickleweed).

## Data analysis Plan

First, with steps 1- , I played with both the vegetation and sediment datasets to understand patterns that each of those explanatory variables had with the landscape and abiotic factors, such as distance from the water source ("Distance"), weather factors ("Season"), time factors ("Year" and "Month"), and location ("Site").

After going through the steps, I went through a two sampled t-test to see if there was a relationship between the sediment fluxes, and my hypothesized factor of maximum plant height of Salicornia pacifica versus the location of the measurements ("Site") to see if those explanatory variables were different between the two sites.

I then ran a general linear mixed model with both sites data and sediment deposition versus the maximum plant height. I followed this by running a general linear mixed model for each site with those same explanatory variables to compare the AIC values and find the best fit. The general linear mixed models I believe demonstrate from their values the clearest relationship that the sites have different relationships with the variables due to Corte Madera's much lower value than San Pablo's glmm, despite them both having the same factors accounted for in the model.

```{r, Load necessary packages}

library(ggplot2)
library(tidyverse)
library(Rmisc)
library(rstatix)
library(tidyr)
library(data.table)
library(dplyr)
library(lme4)
library(stringr)

```

```{r, set working directory}

knitr::opts_knit$set(root.dir = 'C:/Users/savan/OneDrive/Documents/Code/GitHub/DepositionSFB/data')

```

```{r, cleaning data}
# load data
dataset_sed <- read.csv("C:/Users/savan/OneDrive/Documents/Code/GitHub/DepositionSFB/data/SFB_sediment_pad_data (project).csv") 

dataset_veg <- read.csv("C:/Users/savan/OneDrive/Documents/Code/GitHub/DepositionSFB/data/SFB_1m_vegetation_data (project).csv") 

sed <- dataset_sed %>% mutate(FluxAvgRep = ifelse(is.na(FluxAvgRep), 0, FluxAvgRep))

sed <- dataset_sed %>%
  select(-Date.removed.from.field, -Date.placed.in.field)

sed <- sed %>% mutate(FluxAvgRep = ifelse(is.na(FluxAvgRep), 0, FluxAvgRep))

#sed$Plot.ID <- paste(sed$Transect, sed$Distance)
#veg$Plot.ID <- paste(veg$Transect, veg$Distance)
sed$Season <- str_replace_all(sed$Season, " ", "")
sed$Distance <- str_replace_all(sed$Distance, " ", "")
sed$Transect <- str_replace_all(sed$Transect, " ", "")
sed$Year <- str_replace_all(sed$Year, " ", "")
sed$Month <- str_replace_all(sed$Month, " ", "")
sed$FluxAvgRep <- str_replace_all(sed$FluxAvgRep, " ", "")
#sed$Plot.ID <- str_replace_all(sed$Plot.ID, " ", "")
sed$FluxAvgRep <-as.numeric(sed$FluxAvgRep)
sed$FluxAvgRep <- round(sed$FluxAvgRep)
sed$Year <- as.character(sed$Year)
sed$Distance <-as.numeric(sed$Distance)


veg <- dataset_veg %>% mutate(Avg.Ht = ifelse(is.na(Avg.Ht), 0, Avg.Ht))
veg <- dataset_veg %>% mutate(Max.Ht = ifelse(is.na(Max.Ht), 0, Max.Ht))
veg$Season <- str_replace_all(veg$Season, " ", "")
veg$Distance <- str_replace_all(veg$Distance, " ", "")
veg$Transect <- str_replace_all(veg$Transect, " ", "")
veg$Year <- str_replace_all(veg$Year, " ", "")
veg$Month <- str_replace_all(veg$Month, " ", "")
veg$Year <- as.character(veg$Year)
veg$Cover <-as.numeric(veg$Cover)
veg$Distance <-as.numeric(veg$Distance)

str(sed)
str(veg)

```

```{r, mean and median and standard deviation of sediment flux}

sed_mean <- mean(sed$FluxAvgRep, na.rm = TRUE)
sed_mean
sed_max  <- max(sed$FluxAvgRep, na.rm = TRUE)
sed_max
sed_min  <- min(sed$FluxAvgRep, na.rm = TRUE)
sed_min
sed_sd <-sd(sed$FluxAvgRep, na.rm =TRUE)
sed_sd

```

```{r, veg and SAPA}

SAPA_time_space <- ggplot(data=veg, aes(x=Distance, y=Max.Ht, color=Site)) + geom_point()
SAPA_time_space

sed_space <-ggplot(data=sed, aes(x=Distance, y=FluxAvgRep, color=Site)) +geom_point() + geom_jitter(width = 0.2, height = 0.2, alpha = 0.6) + theme_minimal() +
 coord_cartesian(ylim = c(0, 357))
sed_space

sed_time <-ggplot(data=sed, aes(x=Season, y=FluxAvgRep, color=Site)) + geom_point() + geom_jitter(width = 0.2, height = 0.2, alpha = 0.6) + theme_minimal() +
 coord_cartesian(ylim = c(0, 357))
sed_time
```

```{r, visualizing the vegetation data}

SAPA_height <- ggplot(veg, aes(x=Max.Ht,fill=Season)) + geom_boxplot()
SAPA_height

SAPA_cover <- ggplot(veg, aes(x=Cover, fill = Season)) + geom_boxplot()
SAPA_cover

SAPA_Site_cover <- ggplot(veg, aes(x= Cover, fill = Site)) + geom_boxplot()
SAPA_Site_cover

SAPA_Site_mxht <- ggplot(veg, aes(x = Max.Ht, fill = Site)) + geom_boxplot()
SAPA_Site_mxht

#check the data for normality, Max.Ht 
SAPA_Maxht_Site_Normality <-ggplot(veg, aes(x=Max.Ht)) + geom_histogram(binwidth = 0.5)
SAPA_Maxht_Site_Normality

#check the data for normality, Avg.ht  
SAPA_Avght_Site_Normality <-ggplot(veg, aes(x=Avg.Ht)) + geom_histogram(binwidth = 0.5)
SAPA_Avght_Site_Normality

#check the data for normality, Cover 
SAPA_Cover_Site_Normality <-ggplot(veg, aes(x=Cover)) + geom_histogram(binwidth = 0.5)
SAPA_Cover_Site_Normality
```

```{r, putting together two different dataframes}

sed_SAPA <- merge(sed, veg, by = c("Site", "Season", "Distance", "Transect", "Year", "Month"), all = FALSE)

#separating the main dataframe into two site sets 

CMsed_veg <- subset(sed_SAPA, Site == "Corte Madera")
#CMsed_veg

SPsed_veg <- subset(sed_SAPA, Site == "San Pablo")
#SPsed_veg
```

```{r, visualizing the sediment data}

sed_year <- ggplot(sed, aes(x= Year, y= FluxAvgRep, color= Site)) +geom_point()
sed_year

sed_Siteseason <- ggplot(sed, aes(x= Site, y=FluxAvgRep, color = Season)) +geom_point() + geom_boxplot()
sed_Siteseason

sed_site <- ggplot(sed, aes(x=Site, y=FluxAvgRep)) +geom_point() + geom_boxplot()
sed_site

sedvegCM <- ggplot(CMsed_veg, aes(x = Distance, y = FluxAvgRep, color = Season)) + geom_point()
sedvegCM

sedvegSP <- ggplot(SPsed_veg, aes(x = Distance, y = FluxAvgRep, color = Season)) + geom_point() 
sedvegSP

sed_site_normality <- ggplot(sed, aes(x=FluxAvgRep)) +geom_histogram(binwidth = 0.5)
sed_site_normality

```

```{r,two sample t-test via Site and max height}

t.test(Max.Ht~Site,alternative ="two.sided", var.equal = TRUE, data=sed_SAPA)
```

```{r, two sample t-test via Site and sediment flux}

t.test(FluxAvgRep~Site,alternative ="two.sided", var.equal = TRUE, data=sed_SAPA)

```

```{r, Fit GLM with Poisson family}

sedvegglm <-glm(FluxAvgRep~Max.Ht, family = "poisson", data=sed_SAPA)
sedvegglm

sedvegglm_AIC_text <- paste("AIC: 3788")
sedvegglm_AIC_text

sedvegglm_coefficient_Max.Ht <- paste("Coefficient for Max Height: -0.003517 ")
sedvegglm_coefficient_Max.Ht
```

```{r, Plot model diagnostics}

plot(sedvegglm)
```

```{r, glmm}
library(glmmTMB)
sed_glmm = glmmTMB(FluxAvgRep~Max.Ht + Distance + Avg.Ht + (1|Season) + (1|Site) +(1|Year), family = poisson, data=sed_SAPA)
summary(sed_glmm)

AIC(sed_glmm)

```

```{r}
sed_glmmSP = glmmTMB(FluxAvgRep~Max.Ht + Distance + (1|Season) + (1|Year), family = poisson, data=SPsed_veg)
summary(sed_glmmSP)

```

```{r}
sed_glmmCM = glmmTMB(FluxAvgRep~Max.Ht + Distance + (1|Season) + (1|Year), family = poisson, data=CMsed_veg)
summary(sed_glmmCM)
```

```{r}
sed1_cor <- cor_test(FluxAvgRep, Max.Ht, data=sed_SAPA)

#correlation coefficient 
r <- sed1_cor$cor

r
```

```{r}
# plotting the interaction
p5 <- ggplot(sed_SAPA,aes(Max.Ht,FluxAvgRep,group=Season, color=Site)) + geom_point(size=2,alpha=.5,position = position_dodge(width=.15)) +
  geom_smooth(method = "lm", se = FALSE) +
  ylab("Sediment Flux") +
  xlab("Maximum Plant Height") +
  theme_classic(14)
p5

```
