---
title: "Distribution Comparison: Ramped vs. Scarped (Claude)"
format: html
editor: visual
---

```{=html}
<!--
Standalone reference document, same family as marsh_edgeID_vs_site_tran
and averaged_vs_replicate. Built in response to XiaoHang's committee
feedback (classic/robust distribution comparison, specific test name
for comparing two distributions) plus John's earlier framing (SCARP vs.
RAMP variability is a finding in itself, not noise).

Runs on BOTH master_tier5 (averaged) and master_tier5_byrep (replicate-
level, if you've generated it via the updated step6) so you can see
whether the averaging-vs-replicate decision changes which test results
you'd report -- same "explicit comparison, not silent choice" pattern
as the other two docs.

Sources the shared conventions/glossary files for consistent
terminology (PRIMARY/SENSITIVITY, Rule 1a/1b/1c).
-->
```

```{r, setup}
#| message: false

source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\setup.R")
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\plotting.R")
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\modeling_conventions (Claude).R")
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\statistical_tools_glossary (Claude).R")

# make_flux_table() (flux_summary.R) already uses moments::skewness()/
# kurtosis() -- confirmed, so this table uses the same package rather
# than e1071, which was dropped from setup.R to remove the earlier
# skewness()/kurtosis() namespace collision between the two packages.

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/savan/OneDrive - San Francisco State University/ProjectData/GitHub/DepositionSFB/data')

master_tier5 <- readr::read_csv(
  file.path(knitr::opts_knit$get("root.dir"), "master_tier5.csv")
) %>%
  filter(Flux > 0, Dist != 48) %>%
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")))

# Optional -- only run if you've generated this via the updated step6.
# master_tier5_byrep <- readr::read_csv(
#   file.path(knitr::opts_knit$get("root.dir"), "master_tier5_byrep.csv")
# ) %>%
#   filter(Flux > 0, Dist != 48) %>%
#   mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")))
```

## What "compare two distributions" actually means -- three different questions

This document runs three separate tests because "are Ramped and Scarped different" is really three distinct questions, and conflating them is the easiest way to over- or under-claim a result at the defense:

1.  **Is the *spread* different?** (Fligner-Killeen) -- this is John's "variation itself is a finding" point, tested directly rather than eyeballed from a boxplot.
2.  **Is the *whole distribution* (shape + spread + location together) different?** (Kolmogorov-Smirnov) -- this is the literal answer to "what's the test for comparing two distributions."
3.  **Is the *typical value* (location/rank) different?** (Mann-Whitney U / Wilcoxon rank-sum) -- your instinct, and the right test for "which one tends to be higher," but see the caveat below on what it assumes.

## Step 1: Visual comparison + shape statistics

```{r, histogram and shape stats}
ggplot(master_tier5, aes(x = Flux, fill = marsh_edge_type)) +
  geom_histogram(aes(y = after_stat(density)), bins = 40, alpha = 0.6,
                  position = "identity") +
  scale_fill_manual(values = marsh_edge_colors) +
  facet_wrap(~ marsh_edge_type, ncol = 1, scales = "free_y") +
  labs(title = "Distribution of Sediment Flux by Edge Type",
       x = expression(Flux~(g/m^2/day)), y = "Density", fill = "Marsh Edge Type") +
  theme_bw() +
  theme(legend.position = "none")

# Shape statistics table, extending the mean/SD/skewness already in
# make_flux_table() (flux_summary.R) with kurtosis -- the piece
# XiaoHang's feedback specifically asked for that wasn't already
# computed anywhere in the pipeline.
shape_stats <- master_tier5 %>%
  group_by(marsh_edge_type) %>%
  summarise(
    n         = n(),
    mean      = mean(Flux, na.rm = TRUE),
    sd        = sd(Flux, na.rm = TRUE),
    median    = median(Flux, na.rm = TRUE),
    skewness  = moments::skewness(Flux, na.rm = TRUE),
    kurtosis  = moments::kurtosis(Flux, na.rm = TRUE),   # RAW kurtosis --
    # moments::kurtosis() does NOT subtract 3, so a normal distribution
    # reads as ~3 here, not 0. (This is the opposite convention from
    # e1071::kurtosis(), which reports EXCESS kurtosis, normal = 0 --
    # if you ever see a kurtosis table with values near 0 instead of
    # near 3, that's a sign it was built with e1071 instead of moments,
    # not that the distribution itself changed.)
    shapiro_p = shapiro.test(Flux)$p.value,
    .groups = "drop"
  )
shape_stats
```

## Step 2: Is the spread different? (Fligner-Killeen)

```{r, fligner killeen}
# Fligner-Killeen is the NON-PARAMETRIC alternative to Levene's test --
# tests equality of variance/spread WITHOUT assuming normality, which
# matters here since Shapiro already rejects normality for both groups
# (shape_stats$shapiro_p above). Given the SD gap already visible in
# shape_stats (Scarped's SD is roughly an order of magnitude larger),
# expect this to come back significant -- report it as a primary
# finding, not a footnote, per John's framing.

fligner.test(Flux ~ marsh_edge_type, data = master_tier5)
```

## Step 3: Is the whole distribution different? (Kolmogorov-Smirnov)

```{r, ks test}
# The literal answer to "what's the test for comparing two
# distributions" -- tests location, spread, and shape jointly, with no
# distributional assumptions. This is the test to cite if XiaoHang or
# the committee asks for one specific name.

flux_ramped  <- master_tier5$Flux[master_tier5$marsh_edge_type == "Ramped"]
flux_scarped <- master_tier5$Flux[master_tier5$marsh_edge_type == "Scarped"]

ks.test(flux_ramped, flux_scarped)
```

## Step 4: Is the typical value different? (Mann-Whitney U)

```{r, mann whitney}
# CAVEAT: Mann-Whitney's "which group tends to be higher" interpretation
# technically assumes similar spread between groups. Given the Fligner-
# Killeen result above almost certainly shows spread IS different, a
# significant result here may partly reflect that spread difference
# rather than a pure location shift -- report alongside Step 2, not
# instead of it, and don't claim "Scarped values are simply higher"
# from this test alone if Step 2 was also significant.

wilcox.test(Flux ~ marsh_edge_type, data = master_tier5)

# Effect size (rank-biserial correlation) -- Mann-Whitney's p-value
# alone doesn't say how big the difference is, same issue as reporting
# only a p-value for a t-test.
# install.packages("rstatix")  # if not already installed
library(rstatix)
master_tier5 %>% wilcox_effsize(Flux ~ marsh_edge_type)
```

```{r, brunner munzel, eval = FALSE}
# OPTIONAL, more robust alternative: the Brunner-Munzel test asks the
# same "which group tends to be higher" question as Mann-Whitney, but
# does NOT assume similar spread between groups -- built for exactly
# the situation Step 2 is likely to reveal here. Worth running as a
# cross-check if Mann-Whitney and Fligner-Killeen are both significant,
# since it isolates the location effect from the spread effect that
# Mann-Whitney alone can't fully separate.
# install.packages("brunnermunzel")
library(brunnermunzel)
brunnermunzel.test(Flux ~ marsh_edge_type, data = master_tier5)
```

## How to report this as one coherent result

Given the likely pattern (spread differs a lot, shape/skewness differs, and there may or may not be a clean location shift once spread is accounted for), the honest multi-test story is:

1.  **Lead with Fligner-Killeen** -- Scarped sites show significantly greater variability in sediment flux than Ramped sites. This is the headline result matching John's "variation itself is a finding" framing.
2.  **Follow with Kolmogorov-Smirnov** -- the two edge types' overall Flux distributions differ significantly (covers XiaoHang's literal ask).
3.  **Report Mann-Whitney/Brunner-Munzel with the spread caveat stated explicitly** -- e.g. "Scarped values also tend to rank higher (Mann-Whitney, p = ...), though given the spread difference above, a Brunner-Munzel test was used to confirm this isn't purely an artifact of unequal variance."

This is also a good complement to the raincloud plot already in RQ1-spatial.Rmd -- that plot shows the shape difference visually; this document is the statistical backing for exactly what it shows.

## A note on how this fits the classic-tests-vs-GLMM question

This document doesn't resolve whether to lead the thesis with these classic two-group tests or with the Gamma GLMMs already built across RQ1-3 -- that's the open tension between John's and XiaoHang's feedback flagged separately. What it does do: these tests work on the raw two-group comparison (`marsh_edge_type` alone) without needing the `Site/Tran` nesting structure, so they're valid regardless of how that larger debate resolves, and they don't assume the Gamma link function the GLMMs rely on. Worth having as a model-independent sanity check either way -- if a GLMM and Mann-Whitney/KS agree, that's a stronger result than either alone; if they disagree, that's worth understanding before the defense, not after.
