---
title: "Research Question 1 (Distance Deep Dive)"
format: docx
editor: visual
---

Renamed and rebuilt from the old marsh_areas.qmd. Explores whether Flux varies by marsh position (Edge/Platform/Upland) and whether that pattern differs by Site, marsh_edge_type, or marsh_orientation -- includes both the exploratory visualizations and the formal nested model comparison. Named specifically (not just "Research Question 1") since master_thesis_regressions.Rmd already covers RQ1 more broadly at the full Dist level; this file is the coarser, 3-zone deep dive.

```{r, load packages and data}
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\setup.R")
# Covers everything the old file's individual library() calls needed
# (ggplot2, tidyverse, lme4, lmerTest, ggeffects, patchwork, ggh4x,
# rcartocolor, scales) -- no separate library() calls needed.

master_tier5 <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\master_tier5.csv")

# CSVs don't preserve factor levels -- re-factor after the disk read.
master_tier5$Site <- factor(master_tier5$Site,
  levels = c("GiantMarsh", "SanPablo", "Buck'sLanding", "CorteMadera"))
master_tier5$marsh_edge_type <- as.factor(master_tier5$marsh_edge_type)
master_tier5$marsh_orientation <- as.factor(master_tier5$marsh_orientation)
master_tier5$Study.Phase <- factor(master_tier5$Study.Phase, levels = paste0("M", 1:12))

# Marsh_Area used directly instead of re-deriving it as "DistGroup" --
# same Edge/Platform/Upland thresholds, already built once in
# master_analysis_cleanup.Rmd. Re-deriving it here under a different
# name with different code (nested ifelse vs. case_when) was pure
# duplication.
master_tier5$Marsh_Area <- factor(master_tier5$Marsh_Area, levels = c("Edge", "Platform", "Upland"))

sed_areas <- master_tier5
```

```{r, distribution by site}
# Two views of the same thing (bar totals, then boxplot distribution) --
# kept both since they show different things (totals vs. spread), but
# flagging: this boxplot is very close in structure to Plot 1 in the
# combined panels below (same Marsh_Area x Site boxplot). Worth
# reviewing whether one of the two is redundant once you see them
# side by side.

ggplot(sed_areas, aes(x = Marsh_Area, y = Flux)) +
  geom_col(position = position_dodge(preserve = "single")) +
  scale_y_log10() +
  facet_grid(Site ~ ., scales = "fixed") +
  labs(
    x = "Marsh Areas",
    y = "Total flux (g/m²/day)",
    title = "Sediment Flux Across Distances From Marsh Edge at Four Marshes"
  ) +
  theme_minimal(base_size = 64) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 78),
    strip.text = element_text(face = "bold", size = 64),
    legend.position = "bottom",
    legend.text = element_text(size = 80),
    legend.title = element_text(size = 88),
    legend.key.size = unit(3, "cm"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 60),
    axis.text.y = element_text(size = 60),
    axis.title = element_text(size = 60)
  )

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\sed_marshareas2.png",width = 60, height = 40, dpi = 300, limitsize = FALSE)

file.info("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\sed_marshareas2.png")$mtime

ggplot(sed_areas, aes(x = Marsh_Area, y = Flux)) +
  geom_boxplot() +
  scale_y_log10() +
  facet_grid(Site ~ ., scales = "fixed") +
  annotation_logticks(sides = "1") +
  labs(
    x = "Marsh Areas",
    y = "Total flux (g/m²/day)",
    title = "Sediment Flux Across Distances From Marsh Edge at Four Marshes"
  ) +
  theme_minimal(base_size = 64) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 78),
    strip.text = element_text(face = "bold", size = 64),
    legend.position = "bottom",
    legend.text = element_text(size = 80),
    legend.title = element_text(size = 88),
    legend.key.size = unit(3, "cm"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 60),
    axis.text.y = element_text(size = 60),
    axis.title = element_text(size = 60)
  )

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\sed_marshareas3.png",
       width = 60, height = 40, dpi = 300, limitsize = FALSE)
```

```{r, combined panel function}
# The three blocks below (by Site, by marsh_edge_type, by
# marsh_orientation) were identical code three times over, differing
# only in the grouping variable and output filename. Consolidated into
# one function instead.

make_marsh_area_panel <- function(data, group_var, panel_title, filename) {
  group_sym <- rlang::sym(group_var)

  summary_df <- data %>%
    dplyr::group_by(!!group_sym, Marsh_Area) %>%
    dplyr::summarise(
      mean_flux = mean(Flux, na.rm = TRUE),
      sd_flux = sd(Flux, na.rm = TRUE),
      n = length(Flux),
      .groups = "drop"
    ) %>%
    dplyr::mutate(se_flux = sd_flux / sqrt(n))

  p1 <- ggplot(data, aes(x = Marsh_Area, y = Flux)) +
    geom_boxplot(aes(fill = Marsh_Area), outlier.shape = 16, outlier.size = 3) +
    scale_y_log10() +
    facet_grid(as.formula(paste(group_var, "~ .")), scales = "fixed") +
    annotation_logticks(sides = "1") +
    labs(x = "Marsh Area", y = "Sediment Flux (g/m²/day)",
         title = "A) Distribution of Flux by Marsh Area") +
    theme_bw(base_size = 64) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 70),
      strip.text = element_text(face = "bold", size = 64),
      legend.position = "none",
      axis.text.x = element_text(angle = 45, hjust = 1, size = 60),
      axis.text.y = element_text(size = 60),
      axis.title = element_text(size = 60)
    )

  p2 <- ggplot(summary_df, aes(x = Marsh_Area, y = mean_flux,
                                colour = !!group_sym, group = !!group_sym, fill = !!group_sym,
                                ymin = mean_flux - se_flux, ymax = mean_flux + se_flux)) +
    geom_ribbon(alpha = 0.15, colour = NA) +
    geom_line(linewidth = 1.5) +
    geom_point(size = 5) +
    scale_y_log10() +
    labs(x = "Marsh Area", y = "Mean Flux — log scale (g/m²/day)",
         title = "B) Distance Decay Across Sites") +
    theme_bw(base_size = 64) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 70),
      legend.position = "bottom",
      legend.text = element_text(size = 60),
      legend.title = element_text(size = 64),
      legend.key.size = unit(3, "cm"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 60),
      axis.text.y = element_text(size = 60),
      axis.title = element_text(size = 60)
    )

  combined <- p1 + p2 +
    plot_annotation(
      title = panel_title,
      theme = theme(plot.title = element_text(hjust = 0.5, size = 78))
    )

  ggsave(file.path("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images", filename),
         plot = combined, width = 80, height = 45, dpi = 300, limitsize = FALSE)

  combined
}
```

```{r, panel by site}
make_marsh_area_panel(sed_areas, "Site",
  "Sediment Flux Across Distances From Marsh Edge at Four Marshes",
  "sed_combined_panel.png")
```

```{r, panel by marsh edge type}
make_marsh_area_panel(sed_areas, "marsh_edge_type",
  "Sediment Flux Across Distances From Marsh Edge at Four Marshes",
  "sed_combined_panel2.png")
```

```{r, panel by marsh orientation}
make_marsh_area_panel(sed_areas, "marsh_orientation",
  "Sediment Flux Across Distances From Marsh Edge at Four Marshes",
  "sed_combined_panel3.png")
```

```{r, statistics}
# Season replaced -- doesn't exist anymore. m3 already used Study.Phase
# alongside Season, so Study.Phase is this file's own established
# temporal variable, not something new being introduced. Extended the
# same choice to marsh_area1, which only had Season and would
# otherwise lose its temporal control entirely.

sed_areas_clean <- sed_areas %>% filter(Flux > 0)

marsh_area1 <- lmer(log(Flux) ~ Marsh_Area + marsh_edge_type + Study.Phase + (1|Site), data = sed_areas_clean)
summary(marsh_area1)

# Null
m0 <- lmer(log(Flux) ~ 1 + (1|Site), data = sed_areas_clean)

# Spatial structure only
m1 <- lmer(log(Flux) ~ Marsh_Area + (1|Site), data = sed_areas_clean)

# Add edge type interaction
m2 <- lmer(log(Flux) ~ Marsh_Area * marsh_edge_type + (1|Site), data = sed_areas_clean)

# Full model
m3 <- lmer(log(Flux) ~ Marsh_Area * marsh_edge_type + Study.Phase + (1|Site),
           data = sed_areas_clean)

# Compare
AIC(m0, m1, m2, m3)
```
