---
title: "Research Question 2 (USGS vs. Thesis)"
format: docx
editor: visual
---

Repurposed from the old Research_question_2.qmd per Session_Status step 5. Loads usgs_thesis.rds (built by usgs_thesis_merge.qmd) instead of relying on session-leftover objects. Two model variants throughout -- Month and Phen_Period -- replacing Study.Phase and Season, neither of which exist in the merged data anymore.

Study.Dataset (USGS vs Thesis) replaces Year as the actual comparison variable throughout every model and plot. Year had 4 messy, non-adjacent levels (2022, 2023, 2025, 2026) that only ever served as a noisy proxy for which dataset a row came from -- Study.Dataset says that directly. Also fixes a real bug this surfaced: Year was never re-factored after the CSV round-trip in usgs_thesis_merge.qmd, so it silently came through as continuous numeric, which is why an earlier version of this file's emmeans plot rendered Year as a blue color gradient instead of distinct groups.

```{=html}
<!--
Consolidated this session: a standalone "distance visualizations" file
was drafted alongside this one, but its content either duplicated RQ1
(raw-Dist plots, already covered by RQ1-spatial.Rmd and
Research_question_1_distance_deepdive.qmd) or was USGS-only (not an
actual USGS-vs-Thesis comparison, which is RQ2's real question). Folded
in as one new section instead: Marsh Area x Study.Dataset, faceted by
Site -- genuinely new, since the existing rq2 boxplot below only breaks
the USGS-vs-Thesis comparison down by Site, not by Marsh_Area within
each site. Uses Month/Phen_Period (not Study.Phase), matching the
temporal variables already proven to work on thesis_rq2 in this file --
no new data-source guessing needed.
-->
```

```{r, load packages}
library(ggplot2)
library(tidyverse)
library(rstatix)
library(tidyr)
library(data.table)
library(dplyr)
library(lme4)
library(stringr)
library(readxl)
library(lmerTest)
library(ggeffects)
library(ggforce)
library(emmeans)   # was missing -- needed for the emmeans() calls below
library(rcartocolor)
library(scales)

knitr::opts_knit$set(root.dir = 'C:/Users/savan/OneDrive - San Francisco State University/ProjectData/GitHub/DepositionSFB/data')
```

```{r, load merged dataset}
usgs_thesis <- readRDS("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis.rds")

str(usgs_thesis)
```

```{r, build thesis_rq2}
# Despite the name, this keeps BOTH datasets -- it only drops the two
# thesis-only sites that have no USGS counterpart (Buck'sLanding,
# GiantMarsh), so Corte Madera + San Pablo rows from both USGS and
# Thesis remain.
#
# Restricted to Marsh_Area %in% c("Edge", "Platform") -- excludes
# Upland everywhere (thesis Dist=48, USGS Corte Madera Dist=48/30,
# USGS San Pablo Dist=72/150) and Intertidal at San Pablo (Dist 0.5
# through 18). This is the correct way to do this: using Marsh_Area
# instead of raw Dist values handles the fact that Dist=48 means
# Platform at San Pablo but Upland at Corte Madera -- filtering by a
# raw Dist number instead would have either wrongly kept or wrongly
# dropped one of those two, depending which way it was done.

thesis_rq2 <- usgs_thesis %>%
  filter(!Site %in% c("Buck'sLanding", "GiantMarsh")) %>%
  filter(Marsh_Area %in% c("Edge", "Platform"))

thesis_rq2 <- thesis_rq2 %>%
  select(-marsh_edge_type, -marsh_orientation)

thesis_rq2$Study.Dataset <- as.factor(thesis_rq2$Study.Dataset)

unique(thesis_rq2$Site)
unique(thesis_rq2$Study.Dataset)
unique(thesis_rq2$Marsh_Area)  # should show only Edge, Platform

str(thesis_rq2)
nrow(thesis_rq2)
```

```{r, statistics}
shapiro.test(thesis_rq2$Flux)

qqnorm(thesis_rq2$Flux)
qqline(thesis_rq2$Flux)
```

## Model 1: Study.Dataset x temporal grouping, random intercept for Site

```{r, model 1 - Month variant}
usgsthesis_model_month <- lmer(
  Flux ~ Study.Dataset + Month + Study.Dataset * Month + Site,
  data = thesis_rq2
)
summary(usgsthesis_model_month)
```

```{r, model 1 - Phen_Period variant}
usgsthesis_model_phenperiod <- lmer(
  Flux ~ Study.Dataset + Phen_Period + Study.Dataset * Phen_Period + (1 | Site),
  data = thesis_rq2
)
summary(usgsthesis_model_phenperiod)
```

## Model 2: Site effect, random intercepts for Study.Dataset + temporal grouping

```{r, model 2 - Month variant}
usgsthesis_model2_month <- lmer(
  Flux ~ Site + (1 | Study.Dataset) + (1 | Month),
  data = thesis_rq2
)
summary(usgsthesis_model2_month)
```

```{r, model 2 - Phen_Period variant}
usgsthesis_model2_phenperiod <- lmer(
  Flux ~ Site + (1 | Study.Dataset) + (1 | Phen_Period),
  data = thesis_rq2
)
summary(usgsthesis_model2_phenperiod)
```

## Estimated marginal means -- Month variant

```{r, emmeans and plots - Month variant}
emm_month <- emmeans(usgsthesis_model_month, ~ Month * Study.Dataset)
emm_month_df <- as.data.frame(emm_month)

ggplot(emm_month_df, aes(x = Month, y = emmean, color = Study.Dataset, group = Study.Dataset)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.2, position = position_dodge(0.3)) +
  geom_line(position = position_dodge(0.3)) +
  labs(title = "Estimated Marginal Means by Month and Study Dataset",
       y = "log(Flux)", x = "Month", color = "Study Dataset") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(thesis_rq2, aes(x = Site, y = Flux, fill = Study.Dataset)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(aes(color = Study.Dataset),
              alpha = 0.4, size = 0.8, width = 0.2) +
  labs(title = str_wrap("Research Question 2: Is Sediment Deposition Consistent Between USGS and Thesis Study Periods at Corte Madera Marsh and San Pablo NWR?", width = 40),
       y = "Log of Average Flux (g/m^3/day)",
       x = "Site",
       fill = "Study Dataset", color = "Study Dataset") +
  scale_y_log10() +
  theme_bw() +
  theme(
    axis.title.y = element_text(size = 13),
    axis.text.y = element_text(size = 13),
    axis.text.x = element_text(size = 13),
    axis.title.x = element_text(size = 13),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    plot.title = element_text(hjust = 0.5)
  )
```

## Estimated marginal means -- Phen_Period variant

```{r, emmeans and plots - Phen_Period variant}
emm_phenperiod <- emmeans(usgsthesis_model_phenperiod, ~ Phen_Period * Study.Dataset)
emm_phenperiod_df <- as.data.frame(emm_phenperiod)

ggplot(emm_phenperiod_df, aes(x = Phen_Period, y = emmean, color = Study.Dataset, group = Study.Dataset)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.2, position = position_dodge(0.3)) +
  geom_line(position = position_dodge(0.3)) +
  labs(title = "Estimated Marginal Means by Phenological Period and Study Dataset",
       y = "log(Flux)", x = "Phenological Period", color = "Study Dataset") +
  theme_bw()
```

## Marsh Area comparison (USGS vs Thesis)

New this session -- the boxplot above (in the Month-variant chunk) shows Flux by Site x Study.Dataset, but doesn't break that down by Marsh_Area within each site. This does, in both Month and Phen_Period variants to match the rest of the file.

```{r, marsh area comparison - Month variant}
# REMEMBER: this runs on thesis_rq2, so it inherits the same
# restrictions as the rest of the file -- only Corte Madera + San Pablo
# (no Buck'sLanding/GiantMarsh), and only Edge/Platform (no Upland). If
# you ever want this comparison to include Upland or the other two
# sites, build a separate object without those filters rather than
# reusing thesis_rq2 here.

marsh_area_month <- thesis_rq2 %>%
  group_by(Site, Marsh_Area, Study.Dataset, Month) %>%
  summarise(
    mean_flux = mean(Flux, na.rm = TRUE),
    se_flux = sd(Flux, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(thesis_rq2, aes(x = Marsh_Area, y = Flux, fill = Study.Dataset)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(aes(color = Study.Dataset), alpha = 0.4, size = 0.8,
              position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75)) +
  facet_wrap(~ Site) +
  scale_y_log10() +
  labs(
    title = "Sediment Flux by Marsh Area, Site, and Study Dataset",
    y = "Log of Flux (g/m²/day)", x = "Marsh Area",
    fill = "Study Dataset", color = "Study Dataset"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

ggplot(marsh_area_month, aes(x = Month, y = mean_flux, color = Study.Dataset, group = Study.Dataset)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_grid(Site ~ Marsh_Area) +
  labs(
    title = "Mean Flux by Marsh Area Across Months, USGS vs Thesis",
    y = expression(Mean~Flux~(g/m^2/day)), x = "Month", color = "Study Dataset"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

```{r, marsh area comparison - Phen_Period variant}
marsh_area_phenperiod <- thesis_rq2 %>%
  group_by(Site, Marsh_Area, Study.Dataset, Phen_Period) %>%
  summarise(
    mean_flux = mean(Flux, na.rm = TRUE),
    se_flux = sd(Flux, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(marsh_area_phenperiod, aes(x = Phen_Period, y = mean_flux, color = Study.Dataset, group = Study.Dataset)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_grid(Site ~ Marsh_Area) +
  labs(
    title = "Mean Flux by Marsh Area Across Phenological Periods, USGS vs Thesis",
    y = expression(Mean~Flux~(g/m^2/day)), x = "Phenological Period", color = "Study Dataset"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

## Bay vs Channel comparison (USGS only)

New section, different question from the rest of this file. Everything
above is USGS-vs-Thesis; this is USGS-internal -- thesis has no
channel-adjacent transects to compare against, so the only place this
comparison can happen is within USGS's own Corte Madera and San Pablo
data, holding Site and Study.Phase window fixed and varying only
Location (bay-facing Tran A/B vs channel-facing Tran C/D).

Reads usgs_bay.csv and usgs_river.csv directly, not usgs_thesis.rds --
that merge only ever included Bay rows (see usgs_thesis_merge.qmd),
so Channel data was never part of it and isn't reachable through
thesis_rq2. Both source files already carry Marsh_Area, Study.Phase,
and (as of the elevation-join pipeline update) Elevation_m/z_star from
master_analysis_usgs.Rmd -- none of that is re-derived here, only
Month (same derivation as usgs_thesis_merge.qmd, since these are the
pre-merge files and don't have it yet) and the new Location label.

CAUTION: same shape as this file's existing (1|Site) singular-fit issue
above -- only 2 sites, so Site x Location is a de facto 2x2 design.
Site is kept as a FIXED effect in the Month-variant model below
(matching Model 1's own Month-variant convention earlier in this file),
not random, for the same reason.

```{r, build bay_channel}
study_month_levels <- c("July", "August", "September", "October",
                         "November", "December", "January", "February",
                         "March", "April", "May", "June")

sed_usgs_bay_rq2 <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_bay.csv") %>%
  mutate(Location = "Bay")

sed_usgs_river_rq2 <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_river.csv") %>%
  mutate(Location = "Channel")

bay_channel <- bind_rows(sed_usgs_bay_rq2, sed_usgs_river_rq2)

# Re-factor after the CSV round-trip -- same necessity as
# usgs_thesis_merge.qmd's equivalent chunk, not duplicated logic.
bay_channel$Location    <- factor(bay_channel$Location, levels = c("Bay", "Channel"))
bay_channel$Study.Phase <- factor(bay_channel$Study.Phase, levels = paste0("M", 1:12))
bay_channel$Marsh_Area  <- factor(bay_channel$Marsh_Area, levels = c("Intertidal", "Edge", "Platform", "Upland"))
bay_channel$Site        <- as.factor(bay_channel$Site)

# Month wasn't carried into usgs_bay.csv/usgs_river.csv (only added
# during the merge step) -- derive it fresh here, same recipe as
# usgs_thesis_merge.qmd.
bay_channel <- bay_channel %>%
  mutate(Month = factor(month(Date.removed.from.field, label = TRUE, abbr = FALSE),
                         levels = study_month_levels, ordered = TRUE))

# Sanity checks before modeling
unique(bay_channel$Location)
unique(bay_channel$Site)
table(bay_channel$Site, bay_channel$Location)
nrow(sed_usgs_bay_rq2) + nrow(sed_usgs_river_rq2) == nrow(bay_channel)
```

```{r, bay channel model - Month variant}
bay_channel_model_month <- lmer(
  Flux ~ Location + Month + Location * Month + Site,
  data = bay_channel
)
summary(bay_channel_model_month)
```

```{r, bay channel model - Phen_Period variant}
bay_channel_model_phenperiod <- lmer(
  Flux ~ Location + Phen_Period + Location * Phen_Period + (1 | Site),
  data = bay_channel
)
summary(bay_channel_model_phenperiod)
```

```{r, bay channel emmeans and plots - Month variant}
emm_bay_channel_month <- emmeans(bay_channel_model_month, ~ Month * Location)
emm_bay_channel_month_df <- as.data.frame(emm_bay_channel_month)

ggplot(emm_bay_channel_month_df, aes(x = Month, y = emmean, color = Location, group = Location)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.2, position = position_dodge(0.3)) +
  geom_line(position = position_dodge(0.3)) +
  facet_wrap(~ Site) +
  labs(title = "Estimated Marginal Means by Month and Location (Bay vs Channel)",
       y = "log(Flux)", x = "Month", color = "Location") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(bay_channel, aes(x = Marsh_Area, y = Flux, fill = Location)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(aes(color = Location), alpha = 0.4, size = 0.8,
              position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75)) +
  facet_wrap(~ Site) +
  scale_y_log10() +
  labs(
    title = "Sediment Flux by Marsh Area, Site, and Location",
    y = "Log of Flux (g/m²/day)", x = "Marsh Area",
    fill = "Location", color = "Location"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

```{r, bay channel emmeans and plots - Phen_Period variant}
emm_bay_channel_phenperiod <- emmeans(bay_channel_model_phenperiod, ~ Phen_Period * Location)
emm_bay_channel_phenperiod_df <- as.data.frame(emm_bay_channel_phenperiod)

ggplot(emm_bay_channel_phenperiod_df, aes(x = Phen_Period, y = emmean, color = Location, group = Location)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.2, position = position_dodge(0.3)) +
  geom_line(position = position_dodge(0.3)) +
  facet_wrap(~ Site) +
  labs(title = "Estimated Marginal Means by Phenological Period and Location (Bay vs Channel)",
       y = "log(Flux)", x = "Phenological Period", color = "Location") +
  theme_bw()
```

### Observational comparison to thesis (Edge/Platform only)

A separate, restricted object -- not used in any model above, and not
itself a statistical comparison. thesis_rq2 above is restricted to
Marsh_Area %in% c("Edge", "Platform") to stay comparable to what thesis
actually measured; this applies that same restriction to bay_channel
so the resulting plots can be eyeballed side by side against the
"Marsh Area comparison (USGS vs Thesis)" plots earlier in this file --
same Site x Marsh_Area facet layout, same restriction, different
grouping variable (Location instead of Study.Dataset). This is a visual
reference only: Bay vs Channel and USGS vs Thesis are different
questions on different samples, and this doesn't claim otherwise.

The full bay_channel object (all four Marsh_Area levels) from above is
untouched and still available for the modeling already done.

```{r, bay_channel restricted to Edge/Platform}
bay_channel_edge_platform <- bay_channel %>%
  filter(Marsh_Area %in% c("Edge", "Platform"))

unique(bay_channel_edge_platform$Marsh_Area)
nrow(bay_channel_edge_platform)
```

```{r, bay channel plots restricted to Edge/Platform}
ggplot(bay_channel_edge_platform, aes(x = Marsh_Area, y = Flux, fill = Location)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(aes(color = Location), alpha = 0.4, size = 0.8,
              position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75)) +
  facet_wrap(~ Site) +
  scale_y_log10() +
  labs(
    title = "Sediment Flux by Marsh Area (Edge/Platform), Site, and Location",
    subtitle = "Restricted to match the Marsh Area comparison (USGS vs Thesis) plots above -- visual reference, not a statistical comparison",
    y = "Log of Flux (g/m²/day)", x = "Marsh Area",
    fill = "Location", color = "Location"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 10, color = "gray40"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

bay_channel_ep_month <- bay_channel_edge_platform %>%
  group_by(Site, Marsh_Area, Location, Month) %>%
  summarise(
    mean_flux = mean(Flux, na.rm = TRUE),
    se_flux = sd(Flux, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(bay_channel_ep_month, aes(x = Month, y = mean_flux, color = Location, group = Location)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_grid(Site ~ Marsh_Area) +
  labs(
    title = "Mean Flux by Marsh Area (Edge/Platform) Across Months, Bay vs Channel",
    subtitle = "Same facet layout as the Marsh Area (USGS vs Thesis) Month plot above",
    y = expression(Mean~Flux~(g/m^2/day)), x = "Month", color = "Location"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.subtitle = element_text(hjust = 0.5, size = 10, color = "gray40"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

bay_channel_ep_phenperiod <- bay_channel_edge_platform %>%
  group_by(Site, Marsh_Area, Location, Phen_Period) %>%
  summarise(
    mean_flux = mean(Flux, na.rm = TRUE),
    se_flux = sd(Flux, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(bay_channel_ep_phenperiod, aes(x = Phen_Period, y = mean_flux, color = Location, group = Location)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_grid(Site ~ Marsh_Area) +
  labs(
    title = "Mean Flux by Marsh Area (Edge/Platform) Across Phenological Periods, Bay vs Channel",
    subtitle = "Same facet layout as the Marsh Area (USGS vs Thesis) Phen_Period plot above",
    y = expression(Mean~Flux~(g/m^2/day)), x = "Phenological Period", color = "Location"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.subtitle = element_text(hjust = 0.5, size = 10, color = "gray40"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```

```{r, transect comparison by site}
# NOTE: the original file used an object called usgsthesiscomp here,
# which is never built anywhere in this document or its predecessor --
# a fully broken reference, not a session-dependency issue. Using
# thesis_rq2 instead, which is the closest existing equivalent. Flag
# if usgsthesiscomp was meant to be something more specific (e.g. a
# transect-level aggregate) and I'll rebuild that instead.

ggplot(thesis_rq2, aes(x = Site, y = log(Flux), fill = as.factor(Tran), colour = as.factor(Tran))) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA) +
  geom_sina(alpha = 0.6, size = 2) +
  labs(
    x = "Site",
    y = "Log of Average Flux (g/cm³)",
    fill = "Transect",
    colour = "Transect"
  ) +
  theme_classic(base_size = 18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
    strip.background = element_blank(),
    strip.text = element_text(size = 18, face = "bold")
  )
```
