---
title: "RQ1 - Spatial: Averaged-tier analysis"
author: "Savannah K. Miller"
date: "2026-06-22"
output: html_document
---

<!--
SPLIT FILE (averaged tier). RQ1-spatial was split into three files so
each tier can be knit on its own:
  - RQ1-spatial_byrep.Rmd: every analysis on the replicate-level tier
  - RQ1-spatial_avg.Rmd (this file): the same analyses on the averaged
    tier (the chunks that were labeled "ADDED averaged companion")
  - RQ1-spatial_tables.Rmd: the GLMM output-table images, which need
    both tiers. It reads the models the two tier files save at the end,
    so knit both tier files first.
Both tier files keep the same data-loading chunks (loading is fast; the
models are what's slow). If you edit a data-loading chunk, make the
same edit in the other file. Narrative text is kept from the combined
file, so some comments refer to "the byrep chunk above" -- that chunk
now lives in RQ1-spatial_byrep.Rmd.

Split from master_thesis_regressions.Rmd this session into RQ1-spatial.Rmd
(this file) and RQ1-temporal.qmd, so site/edge-type/distance/elevation
structure and Study.Phase/date trends can be read and run separately.
Moved OUT to RQ1-temporal.qmd: the "rq1 time" chunk (rq1_time_trend,
rq1_time -- both by Date.removed.from.field), and two of the three plots
from the "transect x site model with replicate and phase" chunk (the
Study.Phase sina plot and the Dist x Study.Phase facet plot) -- both show
patterns across Study.Phase, not across site/space. The transite1 model
itself and its Tran-by-Site boxplot stay here, since the model's actual
question is the Tran x Site interaction.
-->

```{r, load packages and working directory}

source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\setup.R")
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\plotting.R")
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\flux_summary.R")

# New: the fixed-vs-random-effects rule, the Gamma-primary/Gaussian-
# sensitivity decision, and the model-naming convention used throughout
# this file are now written down ONCE in modeling_conventions (Claude).R instead
# of being re-decided per chunk. Read that file's header comment before
# editing any model below.
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\modeling_conventions (Claude).R")

# Definitions for every statistical function used in this file --
# see the "Statistical Tools Used in This File" section right below,
# or statistical_tools_glossary (Claude).R for the fuller version.
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\statistical_tools_glossary (Claude).R")

# FIX (render): every package this file uses is loaded here, at the top.
# Knitting runs the file top to bottom in a fresh R session, so a
# library() call placed after the first use of its functions fails on
# knit even though it works interactively. Two did that here:
#   - glmmTMB: first used in the "glmm version" chunk, but
#     library(glmmTMB) was ~1,300 lines later ("glmmTMB library" chunk)
#   - DHARMa: used from the "glmm version" chunk on, never loaded
# (both only worked if setup.R or an earlier session had loaded them).
# The rest are loaded before first use already; they're listed here too
# so the file doesn't depend on chunk order. lmerTest loads after lme4 so
# every lmer() gets Satterthwaite p-values -- before, the first lmer()
# fit ran between library(lme4) and library(lmerTest) and got none.
# Loading a package twice is harmless, so the later library() calls are
# left in place.
library(tidyverse)   # dplyr, ggplot2, tidyr, stringr, readr
library(scales)      # scales:: label helpers
library(lme4)        # lmer(), glmer()
library(lmerTest)    # Satterthwaite p-values for lmer()
library(glmmTMB)     # Gamma GLMMs
library(DHARMa)      # residual diagnostics
library(car)         # Anova()
library(emmeans)     # emmeans(), emtrends()
library(ggeffects)   # ggpredict()
library(MuMIn)       # r.squaredGLMM()
library(moments)     # skewness(), kurtosis()
library(ggdist)      # stat_halfeye()
library(tidytext)    # reorder_within()
library(ggbeeswarm)  # geom_quasirandom()
library(patchwork)   # combining plots
library(ggforce)     # geom_sina()
```

<!--
STATISTICAL TOOLS USED IN THIS FILE
(full definitions in statistical_tools_glossary (Claude).R; this is the subset
that actually appears below)

MODEL FITTING
- lmer() [lme4] -- Gaussian linear mixed model; SENSITIVITY check
  version of most models here (Rule 2), usually on log(Flux).
- glmer() [lme4] -- generalized linear mixed model (distmodel only).
- glmmTMB() [glmmTMB] -- PRIMARY model fitter for this project. Set
  family = Gamma(link = "log") for Flux everywhere it's used here.

SIGNIFICANCE / MODEL COMPARISON
- Anova() [car] -- Type II significance test on a fitted model's fixed
  effects. Capital A; different from anova() below.
- anova() [base/lme4] -- either a fixed-effects table for one model, or
  (with two model arguments) a likelihood ratio test comparing two
  NESTED models, e.g. deposition_model_base vs. deposition_model_slope.
- AIC() -- compares non-nested models by fit-vs-complexity tradeoff
  (lower = better); used for distsitemodel_glmm vs. its dispersion/noGM
  variants.

POST-HOC / EFFECT SIZE
- emmeans() / emtrends() [emmeans] -- model-predicted means (emmeans)
  or slopes (emtrends) for specific factor levels/combinations, with
  pairwise comparisons. type = "response" back-transforms Gamma/log-
  link estimates onto the Flux scale.
- ggpredict() [ggeffects] -- model-predicted values across a range of a
  continuous predictor (z_star here), for plotting a fitted line/ribbon.
- r.squaredGLMM() [MuMIn] -- marginal/conditional R² for a mixed model.

DISTRIBUTION / RESIDUAL DIAGNOSTICS
- shapiro.test() / qqnorm() / qqline() -- normality check on raw Flux
  (rejects normality -- part of the Gamma justification, Rule 2). Not
  the right diagnostic for a fitted GLMM's residuals -- that's DHARMa.
- simulateResiduals() / plot() / testDispersion() [DHARMa] -- simulation-
  based residual diagnostics appropriate for a Gamma GLMM, where raw
  residuals never look normal even in a well-fit model. See
  statistical_tools_glossary (Claude).R for how the simulation actually works.

VISUALIZATION
- stat_halfeye() [ggdist] -- draws a half-violin/density ("cloud") of
  the raw distribution; used with geom_boxplot()/geom_jitter() to build
  the raincloud plot comparing Flux distribution shape (skewness)
  between Ramped and Scarped, on the raw (not log-transformed) scale.
-->


```{r, working directory and master data file}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/savan/OneDrive - San Francisco State University/ProjectData/GitHub/DepositionSFB/data')

# 48 m is kept in master_tier5_byrep.csv on purpose (step6 feeds the
# USGS comparison / marsh walk, which needs it), so RQ1 removes it HERE,
# at load. Every later rebuild in this file starts from
# master_tier5_byrep, so 48 m can't come back. The unfiltered copy is
# kept under its own name only for the with/without-48 comparison plot.
master_tier5_byrep_with48 <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\master_tier5_byrep.csv")

master_tier5_byrep <- master_tier5_byrep_with48 |>
  filter(is.na(Dist) | Dist != 48)
stopifnot(!any(master_tier5_byrep$Dist == 48, na.rm = TRUE))

master_tier5_byrep_clean <- master_tier5_byrep |> 
  filter(Flux > 0, 
         Dist != 48)

# FIX: master_tier5_byrep.csv doesn't carry step1's factor level order (CSVs
# only store text, not R's factor metadata) -- without setting this
# explicitly, R/glmmTMB default to alphabetical order, which makes
# Ramped (not Scarped) the reference level, silently flipping every
# model's baseline from what step1 intended. Matches the level order
# set in step1_master_analysis_cleanup.Rmd.
master_tier5_byrep_clean <- master_tier5_byrep_clean %>%
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")))

month_levels <- c("Jul", "Aug", "Sep", "Oct", "Nov", "Dec",
                   "Jan", "Feb", "Mar", "Apr", "May", "Jun")

master_tier5_byrep <- master_tier5_byrep %>%
  mutate(Month = factor(Month, levels = month_levels))

master_tier5_byrep_clean <- master_tier5_byrep_clean %>%
  mutate(Month = factor(Month, levels = month_levels))

# ADDED: averaged companion, same filter/factor steps as master_tier5_byrep
# above, applied to master_tier5.csv (the averaged tier) instead.
# Same as the byrep side: 48 m removed at load, unfiltered copy kept
# only for the with/without-48 plot.
master_tier5_with48 <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\master_tier5.csv")

master_tier5 <- master_tier5_with48 |>
  filter(is.na(Dist) | Dist != 48)
stopifnot(!any(master_tier5$Dist == 48, na.rm = TRUE))

master_tier5_clean <- master_tier5 |> 
  filter(Flux > 0, 
         Dist != 48)

master_tier5_clean <- master_tier5_clean %>%
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")))

master_tier5 <- master_tier5 %>%
  mutate(Month = factor(Month, levels = month_levels))

master_tier5_clean <- master_tier5_clean %>%
  mutate(Month = factor(Month, levels = month_levels))

```

```{r, ADDED consolidated averaged-side data prep}
# This file mutates master_tier5_byrep_clean incrementally throughout
# (Site releveling, Tran/Study.Phase factor conversions, a Flux>0/
# finite-log safety filter before transite1, etc.), scattered across
# many chunks -- several inside plot-only chunks not duplicated here
# (see the note at the end of this file for what wasn't ported).
#
# Rather than replicate every scattered mutation in lockstep, the
# consequential ones (the ones that matter for model FITTING, not just
# plot cosmetics/ordering) are consolidated here, once, applied to
# master_tier5_clean up front. Cosmetic factor-level reordering used
# only by skipped plots (e.g. Site level order for x-axis display) is
# NOT replicated here since it doesn't change model results.
#
# 48 m: removed ONCE, at load in the first chunk, from both
# master_tier5_byrep and master_tier5. Everything in this file is built
# from those two objects, so the byrep rebuild in the z* chunk (and this
# averaged rebuild) can no longer bring 48 m back. The Dist != 48 filters
# still in this file are now harmless no-ops.
#
# Rebuilt averaged tier, swapped in at the same point the byrep side is
# rebuilt (z* chunk), so each _byrep/_avg pair gets identical prep.
# Re-applies the Scarped-first marsh_edge_type order, which rebuilding
# from the raw CSV drops.
master_tier5_clean_rebuilt <- master_tier5 |>
  filter(Flux > 0) |>   # matches the byrep-side rebuild
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")),
         Month = factor(Month, levels = month_levels))

# Matches the Flux>0 / finite-log safety filter applied before transite1
# (originally at line ~1086 of the byrep side) -- redundant given the
# filter just above, but kept for parity/clarity with the original.
master_tier5_clean_rebuilt <- master_tier5_clean_rebuilt[master_tier5_clean_rebuilt$Flux > 0 &
                                     !is.na(master_tier5_clean_rebuilt$Flux) &
                                     is.finite(log(master_tier5_clean_rebuilt$Flux)), ]

# Matches the type/level conversions applied before transite1 (Tran,
# Site as factors; Study.Phase leveled M1-M12)
master_tier5_clean_rebuilt$Tran <- as.factor(master_tier5_clean_rebuilt$Tran)
master_tier5_clean_rebuilt$Site <- as.factor(master_tier5_clean_rebuilt$Site)
master_tier5_clean_rebuilt$Study.Phase <- factor(master_tier5_clean_rebuilt$Study.Phase, levels = paste0("M", 1:12))

# avgdist_avg: averaged companion to avgdist (itself just a filtered
# view, not actually averaged despite the name -- see the "distance
# model prep" chunk's own comment on the byrep side). Re-applies
# Dist != 48 (a no-op now -- removed at load) on top of the rebuilt tier, same as avgdist does on
# master_tier5_byrep_clean.
avgdist_avg <- master_tier5_clean_rebuilt %>%
  filter(Dist != 48) %>%
  mutate(Dist = factor(Dist, levels = c(0.5, 2, 6, 12, 18, 24)))

# avgtotal_avg: averaged companion to avgtotal (a further aggregate()
# on top of whichever base it's built from)
avgtotal_avg <- aggregate(Flux ~ Site + Dist + Study.Phase +
                        marsh_orientation + marsh_edge_type,
                      data = master_tier5_clean_rebuilt, FUN = mean)

# avgdist_protected_avg: averaged companion to avgdist_protected
avgdist_protected_avg <- avgdist_avg %>%
  filter(!Site %in% c("SanPablo", "GiantMarsh"))

# marsh_group_flux_avg: averaged companion to marsh_group_flux
marsh_group_flux_avg <- master_tier5_clean_rebuilt %>% 
  mutate(Marsh_Group = case_when(
    Site %in% c("SanPablo", "GiantMarsh") ~ "OpenBay", 
    Site %in% c("CorteMadera", "Buck'sLanding") ~ "Protected", 
    TRUE ~ NA_character_
  ))
```

## Site / edge type


```{r, ADDED averaged companion for rq1_site}

site_order_avg <- master_tier5 %>%
  distinct(Site, marsh_edge_type) %>%
  arrange(marsh_edge_type, Site) %>%
  pull(Site)

master_tier5_clean <- master_tier5_clean %>%
  mutate(
    Site = factor(Site, levels = site_order_avg), 
    slope_fitted_elevation_m = as.numeric(slope_fitted_elevation_m)
)

rq1_site_avg = ggplot(master_tier5_clean, aes(x =Site , y = log(Flux),
                                                  fill = marsh_edge_type,
                                                  color = marsh_edge_type)) +
  geom_boxplot(width = 0.4, outlier.shape = NA,
               alpha = 0.3,
               position = position_dodge(0.5)) +
  geom_point(aes(color = marsh_edge_type),
             position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5),
             size = 1.5, alpha = 0.3) +
  scale_color_manual(values = marsh_edge_colors) + 
  scale_fill_manual(values = marsh_edge_colors) +
  ylim(0, 6) +
  labs(y = "Log(Flux (g/m^2/day))", x = "Site",
       fill = "Marsh Edge Type", color = "Marsh Edge Type",
       title = str_wrap("Sediment Deposition Rates Across Sites Between Scarped and Ramped Marsh Edge Sites (Averaged)")) +
  theme_bw() +
  theme(legend.position = "bottom")
rq1_site_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq1_site_avg.png", 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

```


```{r, ADDED averaged companion for the raincloud density plot}

rq1_density_sj_avg = master_tier5_clean %>%
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")),
         log_flux = log10(Flux)) %>%
  ggplot(aes(y = log_flux, fill = marsh_edge_type, color = marsh_edge_type)) +
  geom_density(aes(x = after_stat(count)), orientation = "y",
               bw = "SJ", alpha = 0.5) +
  scale_y_continuous(
    labels = function(x) scales::label_math(10^.x, format = force)(x)
  ) +
  scale_fill_manual(values = marsh_edge_colors) +
  scale_color_manual(values = marsh_edge_colors) +
  facet_wrap(~marsh_edge_type, ncol = 2) +
  labs(x = "Count", y = expression(Flux~(g/m^2/day)),
       title = str_wrap("Distribution Shape of Sediment Flux: Ramped vs. Scarped (Averaged)", width = 40)) +
  theme_bw() +
  theme(legend.position = "none")
rq1_density_sj_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_density_sj_avg.png",
       plot = rq1_density_sj_avg,
       width = 8, height = 5, dpi = 300, limitsize = FALSE)

```


```{r, ADDED averaged companion for rq1_boxplot and rq1_boxplot_bysite}
# Same code as the chunk above, run on master_tier5_clean (averaged tier).
# Intermediate objects (plot_data, extreme_zones, p_scarped, ...) are
# reused/overwritten exactly as the byrep chunk does; only the two final
# figures get _avg names.

# library(diptest)  # commented out: no diptest function (dip.test) is used in this file, and it would stop the knit if not installed

library(tidyverse)
library(ggbeeswarm)  # for geom_quasirandom() — install.packages("ggbeeswarm") if needed

# --------------------------------------------------------------------------
# STEP 1: Prep data + compute the stats we need for annotations (#2, #3)
# --------------------------------------------------------------------------

plot_data <- master_tier5_clean %>%
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")),
         log_flux = log10(Flux))

# Per-group summary: n, IQR bounds, and outlier MAGNITUDE (not just count).
#
# FIX: IQR/whiskers are now computed on log10(Flux), not raw Flux, because
# the plot is DISPLAYED on a log10 scale. Computing Tukey bounds on raw,
# heavily right-skewed Flux made "q1 - 1.5*iqr" go negative -- impossible
# for a flux value -- which silently produced zero lower-tail outliers
# (the "min = NA" bug). Doing the same math in log-space keeps the outlier
# rule consistent with what the reader is actually looking at.
#
# NOTE: outlier COUNT/% still looks similar across groups almost by
# construction (Tukey's rule is relative to each group's own IQR) -- the
# real difference is MAGNITUDE, in both directions, so both tails are
# computed and reported separately.
group_stats <- plot_data %>%
  group_by(marsh_edge_type) %>%
  summarise(
    n         = n(),
    q1_log    = quantile(log_flux, 0.25, na.rm = TRUE),
    q3_log    = quantile(log_flux, 0.75, na.rm = TRUE),
    iqr_log   = q3_log - q1_log,
    lower_log = q1_log - 1.5 * iqr_log,
    upper_log = q3_log + 1.5 * iqr_log,
    lower     = 10 ^ lower_log,   # back-transformed to raw Flux units
    upper     = 10 ^ upper_log,
    .groups   = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    upper_vals   = list(plot_data$Flux[plot_data$marsh_edge_type == marsh_edge_type &
                                        plot_data$Flux > upper]),
    lower_vals   = list(plot_data$Flux[plot_data$marsh_edge_type == marsh_edge_type &
                                        plot_data$Flux < lower]),
    # counts kept PER TAIL (not combined) so each annotation only claims
    # the outliers it's actually describing -- avoids the "57 outliers"
    # reading as if it belongs to the max/upper number alone
    n_upper      = length(upper_vals),
    n_lower      = length(lower_vals),
    pct_upper    = round(100 * n_upper / n, 1),
    pct_lower    = round(100 * n_lower / n, 1),
    max_upper    = if (n_upper > 0) max(upper_vals) else NA_real_,
    min_lower    = if (n_lower > 0) min(lower_vals) else NA_real_,
    # fold-multiple beyond the whisker -- the "how far", not "how many"
    upper_ratio  = round(max_upper / upper, 1),
    lower_ratio  = round(lower / min_lower, 1),
    label_top    = str_wrap(paste0(n_upper, " upper outliers (", pct_upper, "%) · max = ",
                           scales::label_comma()(round(max_upper)),
                           " (", upper_ratio, "x whisker)"), width = 26),
    label_bottom = str_wrap(if (n_lower > 0) {
      paste0(n_lower, " lower outliers (", pct_lower, "%) · min = ",
             signif(min_lower, 2), " (", lower_ratio, "x below whisker)")
    } else {
      "no lower-tail outliers"
    }, width = 26)
  ) %>%
  ungroup()

# #3: build "Scarped\n(n = 210)" style axis labels
n_labels <- group_stats %>%
  mutate(axis_label = paste0(marsh_edge_type, "\n(n = ", n, ")")) %>%
  select(marsh_edge_type, axis_label) %>%
  deframe()

# Flag each point as outlier or not, for coloring in #6 (uses the corrected,
# log-space-derived lower/upper bounds)
plot_data <- plot_data %>%
  left_join(group_stats %>% select(marsh_edge_type, lower, upper), by = "marsh_edge_type") %>%
  mutate(is_outlier = Flux < lower | Flux > upper)

# --------------------------------------------------------------------------
# STEP 2: Build the plot
# --------------------------------------------------------------------------
# De-emphasizing the middle: the box/median/IQR are nearly identical between
# groups (confirmed, not just a plotting artifact) -- that's a real finding
# ("central tendency doesn't differ"), but it shouldn't dominate the plot's
# visual weight when the actual story is in the tails. Changes below:
#   - violin now trim = TRUE, so it stops at the real data range instead of
#     a kernel-density tail that visually exaggerates the extremes
#   - violin width/alpha reduced so it reads as context, not the focal shape
#   - light shaded rectangles mark the "beyond whisker" zone in each tail,
#     pulling the eye toward where the groups actually diverge

annot_y_top    <- max(plot_data$Flux, na.rm = TRUE) * 2.2
annot_y_bottom <- min(plot_data$Flux, na.rm = TRUE) * 0.35

# Shaded rectangles highlighting the extreme zones (beyond each group's
# whisker), drawn first so everything else layers on top
extreme_zones <- group_stats %>%
  mutate(
    # constant, not level-dependent: once faceted, each panel centers its
    # one category at x = 1 regardless of the original factor level
    xmin = 1 - 0.35,
    xmax = 1 + 0.35
  )

rq1_boxplot_avg <- ggplot(plot_data, aes(x = marsh_edge_type, y = Flux,
                                      fill = marsh_edge_type, color = marsh_edge_type)) +

  # highlight the upper extreme zone (above whisker) per group
  geom_rect(data = extreme_zones,
            aes(xmin = xmin, xmax = xmax, ymin = upper, ymax = Inf),
            inherit.aes = FALSE, fill = "grey85", alpha = 0.35) +
  # highlight the lower extreme zone (below whisker) per group
  geom_rect(data = extreme_zones,
            aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = lower),
            inherit.aes = FALSE, fill = "grey85", alpha = 0.35) +

  # #1 (revised again): violin made much thinner/subtler -- narrower width,
  # near-invisible fill, and a muted gray outline instead of the group
  # color, so it reads as faint background texture rather than a shape
  # competing with the boxplot/beeswarm/tail story
  geom_violin(fill = "grey60", color = "grey60",
              alpha = 0.06, trim = TRUE, linewidth = 0.25, width = 0.35) +

  geom_boxplot(width = 0.1, alpha = 0.85, outlier.shape = NA, linewidth = 0.5) +

  # #6: quasirandom (beeswarm-style) points instead of default outlier dots
  geom_quasirandom(aes(alpha = is_outlier, size = is_outlier),
                    width = 0.15, shape = 19) +
  scale_alpha_manual(values = c(`FALSE` = 0.12, `TRUE` = 0.9), guide = "none") +
  scale_size_manual(values = c(`FALSE` = 0.8, `TRUE` = 1.9), guide = "none") +

  # #2: magnitude annotations, both tails
  geom_text(data = group_stats,
            aes(x = marsh_edge_type, y = annot_y_top, label = label_top),
            inherit.aes = FALSE, size = 3.1, fontface = "bold", color = "gray20",
            lineheight = 0.9) +
  geom_text(data = group_stats,
            aes(x = marsh_edge_type, y = annot_y_bottom, label = label_bottom),
            inherit.aes = FALSE, size = 3.1, fontface = "bold", color = "gray20",
            lineheight = 0.9) +

  # #5: log scale, explicit note in axis title; extra top/bottom expansion
  # gives the now-wrapped, multi-line annotations room to breathe
  scale_y_log10(labels = scales::label_comma(),
                expand = expansion(mult = c(0.15, 0.35))) +

  # #3: sample size in x-axis labels
  scale_x_discrete(labels = n_labels) +

  # split into one panel per group -- this is the actual fix for the text
  # overlap: each group's annotations now get the FULL panel width instead
  # of sharing one x-axis with the neighboring group's text bleeding in.
  # scales = "free_x" + drop = TRUE means each panel shows only its own
  # single category, centered; y stays fixed/shared so the two panels
  # remain visually comparable.
  facet_wrap(~ marsh_edge_type, nrow = 1, scales = "free_x") +

  scale_fill_manual(values = marsh_edge_colors) +
  scale_color_manual(values = marsh_edge_colors) +

  # #4: title states the finding
  labs(
    x = "Marsh Edge Type",
    y = expression(Flux~(g/m^2/day)~"— log"[10]~"scale"),
    # byrep title states a finding ("Similar Central Flux, but Scarped...") --
    # neutral title here until you've confirmed it also holds when averaged
    title = str_wrap("Flux Distribution by Marsh Edge Type (Averaged)",
                      width = 45),
    subtitle = "Log10 scale; shaded = beyond-whisker zone",
    caption = "Outlier bounds (Tukey's 1.5×IQR) computed on log10(Flux) to match the log-scaled axis."
  ) +

  theme_bw() +
  theme(
    legend.position       = "none",
    plot.title            = element_text(face = "bold", size = 13),
    plot.subtitle         = element_text(size = 9.5, color = "gray30"),
    plot.caption           = element_text(size = 8, color = "gray45", hjust = 0),
    # simplify: no panel box, no vertical gridlines (only 2 categories --
    # nothing for them to line up against), minor gridlines dropped, axis
    # lines kept so the plot still reads as bounded
    panel.border          = element_blank(),
    panel.grid.minor       = element_blank(),
    panel.grid.major.x     = element_blank(),
    axis.line             = element_line(color = "grey40", linewidth = 0.3),
    axis.ticks            = element_line(color = "grey40", linewidth = 0.3),
    plot.margin            = margin(t = 10, r = 14, b = 8, l = 8),
    # facet strip hidden -- the x-axis label (group + n) already identifies
    # each panel, so a strip label would just repeat it
    strip.background       = element_blank(),
    strip.text             = element_blank(),
    panel.spacing           = unit(1.5, "lines")
  )

rq1_boxplot_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_boxplot_avg.png",
       plot = rq1_boxplot_avg,
       width = 8, height = 5.5, dpi = 300, limitsize = FALSE)

library(tidyverse)
library(patchwork)  # combines the Scarped block and Ramped block side by side
                     # install.packages("patchwork") if needed

# --------------------------------------------------------------------------
# STEP 1: Prep data
# --------------------------------------------------------------------------
# ASSUMPTION: your data has a column called `Site` that pairs with the
# `site_colors` named vector, and each Site belongs to exactly one
# marsh_edge_type (e.g. GiantMarsh/CorteMadera -> Scarped;
# SanPablo/Buck'sLanding -> Ramped). If that's not the case, adjust below.

plot_data <- master_tier5_clean %>%
  mutate(marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")),
         Site = factor(Site, levels = names(site_colors)),
         log_flux = log10(Flux))

# --------------------------------------------------------------------------
# STEP 2: Per-SITE Tukey bounds (log space) -- now that each site gets its
# own panel, outlier extremity is assessed against that site's own
# distribution rather than the whole marsh type's.
# --------------------------------------------------------------------------
site_stats <- plot_data %>%
  group_by(marsh_edge_type, Site) %>%
  summarise(
    n         = n(),
    q1_log    = quantile(log_flux, 0.25, na.rm = TRUE),
    q3_log    = quantile(log_flux, 0.75, na.rm = TRUE),
    iqr_log   = q3_log - q1_log,
    lower_log = q1_log - 1.5 * iqr_log,
    upper_log = q3_log + 1.5 * iqr_log,
    lower     = 10 ^ lower_log,
    upper     = 10 ^ upper_log,
    .groups   = "drop"
  ) %>%
  rowwise() %>%
  mutate(
    upper_vals   = list(plot_data$Flux[plot_data$Site == Site & plot_data$Flux > upper]),
    lower_vals   = list(plot_data$Flux[plot_data$Site == Site & plot_data$Flux < lower]),
    n_upper      = length(upper_vals),
    n_lower      = length(lower_vals),
    pct_upper    = round(100 * n_upper / n, 1),
    pct_lower    = round(100 * n_lower / n, 1),
    max_upper    = if (n_upper > 0) max(upper_vals) else NA_real_,
    min_lower    = if (n_lower > 0) min(lower_vals) else NA_real_,
    upper_ratio  = round(max_upper / upper, 1),
    lower_ratio  = round(lower / min_lower, 1),
    label_top    = str_wrap(paste0(n_upper, " upper outliers (", pct_upper, "%) · max = ",
                                    scales::label_comma()(round(max_upper)),
                                    " (", upper_ratio, "x whisker)"), width = 26),
    label_bottom = str_wrap(if (n_lower > 0) {
      paste0(n_lower, " lower outliers (", pct_lower, "%) · min = ",
             signif(min_lower, 2), " (", lower_ratio, "x below whisker)")
    } else {
      "no lower-tail outliers"
    }, width = 26),
    strip_label  = paste0(Site, "\n(n = ", n, ")")
  ) %>%
  ungroup()

plot_data <- plot_data %>%
  left_join(site_stats %>% select(Site, lower, upper), by = "Site") %>%
  mutate(is_outlier = Flux < lower | Flux > upper)

outlier_points <- plot_data %>% filter(is_outlier)

# Shared y-axis range across BOTH blocks so Scarped and Ramped stay visually
# comparable even though they're built as two separate ggplot objects
y_range <- range(plot_data$Flux, na.rm = TRUE)
y_limits <- c(y_range[1] / 9, y_range[2] * 9)

# --------------------------------------------------------------------------
# STEP 3: A function that builds one marsh-edge-type block (its own sites,
# faceted), so Scarped and Ramped are built identically
# --------------------------------------------------------------------------
build_block <- function(edge_type) {

  block_data     <- filter(plot_data, marsh_edge_type == edge_type)
  block_stats    <- filter(site_stats, marsh_edge_type == edge_type)
  block_outliers <- filter(outlier_points, marsh_edge_type == edge_type)
  strip_labels   <- deframe(block_stats %>% select(Site, strip_label))

  annot_y_top    <- y_range[2] * 5
  annot_y_bottom <- y_range[1] / 5

  # NOTE: this is the actual fix -- sites were sharing ONE panel as two
  # x-categories (same crowding bug as before), not truly faceted. Adding
  # facet_wrap here gives each site its own panel at full width, which is
  # also what lets the wrapped annotation text render without overlapping
  # its neighbor.
  ggplot(block_data, aes(x = 1, y = Flux, fill = Site, color = Site)) +

    geom_rect(data = block_stats,
              aes(xmin = 1 - 0.4, xmax = 1 + 0.4, ymin = upper, ymax = Inf),
              inherit.aes = FALSE, fill = "grey85", alpha = 0.35) +
    geom_rect(data = block_stats,
              aes(xmin = 1 - 0.4, xmax = 1 + 0.4, ymin = 0, ymax = lower),
              inherit.aes = FALSE, fill = "grey85", alpha = 0.35) +

    geom_violin(fill = "grey60", color = "grey60",
                alpha = 0.08, trim = TRUE, linewidth = 0.25, width = 0.5) +
    geom_boxplot(width = 0.15, alpha = 0.85, outlier.shape = NA, linewidth = 0.5) +

    geom_jitter(data = block_outliers,
                aes(x = 1),
                position = position_jitter(width = 0.1, height = 0, seed = 1),
                size = 2, shape = 19, alpha = 0.8) +

    geom_text(data = block_stats,
              aes(x = 1, y = annot_y_top, label = label_top),
              inherit.aes = FALSE, size = 2.8, fontface = "bold", color = "gray20",
              lineheight = 0.9) +
    geom_text(data = block_stats,
              aes(x = 1, y = annot_y_bottom, label = label_bottom),
              inherit.aes = FALSE, size = 2.8, fontface = "bold", color = "gray20",
              lineheight = 0.9) +

    scale_fill_manual(values = site_colors, guide = "none") +
    scale_color_manual(values = site_colors) +
    scale_y_log10(labels = scales::label_comma(), limits = y_limits,
                  expand = expansion(mult = c(0.04, 0.08))) +

    facet_wrap(~ Site, nrow = 1, scales = "free_x", labeller = as_labeller(strip_labels)) +

    labs(title = edge_type, x = NULL, y = expression(Flux~(g/m^2/day))) +

    theme_bw() +
    theme(
      legend.position     = "none",   # combined legend added once, below
      plot.title          = element_text(face = "bold", size = 12, hjust = 0.5),
      panel.border        = element_blank(),
      panel.grid.minor    = element_blank(),
      panel.grid.major.x  = element_blank(),
      axis.line           = element_line(color = "grey40", linewidth = 0.3),
      axis.ticks.x        = element_blank(),
      axis.text.x         = element_blank(),
      strip.background    = element_blank(),
      strip.text          = element_text(size = 9, face = "bold"),
      panel.spacing        = unit(1, "lines")
    )
}

# --------------------------------------------------------------------------
# STEP 4: Build both blocks, combine, add one shared legend + overall title
# --------------------------------------------------------------------------
p_scarped <- build_block("Scarped")
p_ramped  <- build_block("Ramped") + theme(axis.text.y = element_blank(),
                                            axis.ticks.y = element_blank(),
                                            axis.line.y = element_blank())

# turn the legend back on for just one panel so patchwork can collect it
p_scarped_legend <- p_scarped + theme(legend.position = "right") +
  guides(color = guide_legend(title = "Site", override.aes = list(size = 3)))

rq1_boxplot_bysite_avg <- (p_scarped_legend | p_ramped) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title    = "Which Sites Drive the Extreme Flux Values? (Averaged)",
    subtitle = "Each panel is one site; grouped by marsh edge type. Log10 scale; shaded = beyond-whisker zone",
    caption  = "Outlier bounds (Tukey's 1.5×IQR) computed on log10(Flux), per site, to match the log-scaled axis.",
    theme = theme(
      plot.title    = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 9.5, color = "gray30"),
      plot.caption  = element_text(size = 8, color = "gray45", hjust = 0)
    )
  )

rq1_boxplot_bysite_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_boxplot_bysite_avg.png",
       plot = rq1_boxplot_bysite_avg,
       width = 11, height = 5.5, dpi = 300, limitsize = FALSE)
```


```{r, ADDED averaged companion for rq1_tran}

tran_order_avg <- master_tier5_clean %>%
  distinct(Tran, marsh_edge_type) %>%
  arrange(marsh_edge_type, Tran) %>%
  pull(Tran)

rq1_tran_avg = ggplot(master_tier5_clean, aes(x =Site , y = log(Flux),
                                    fill = Tran,
                                    color = Tran)) +
  geom_boxplot(width = 0.4, outlier.shape = NA,
               alpha = 0.3,
               position = position_dodge(0.5)) + facet_wrap(~marsh_edge_type, scales = "free_x") +
  geom_point(aes(color = marsh_edge_type),
             position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.5),
             size = 1.5, alpha = 0.3) +
 scale_color_manual(values = site_colors) + 
   scale_fill_manual(values = transect_colors) + 
  scale_x_reordered() +
  ylim(0, 6) +
  labs(y = "Log(Flux (g/m^2/day))", x = "Site",
       fill = "Marsh Edge Type", color = "Marsh Edge Type",
       title = str_wrap("Sediment Deposition Rates Across Transects Between Scarped and Ramped Marsh Edge Sites (Averaged)", width = 40)) +
  theme_bw() +
  theme(legend.position = "bottom")
rq1_tran_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq1_tran_avg.png", 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

```


```{r, ADDED averaged companion for the sample-size bar charts}

edge_counts_by_site_avg <- master_tier5_clean %>%
  count(marsh_edge_type, Site, name = "n")

rq1_edge_counts_site_avg = ggplot(edge_counts_by_site_avg,
                               aes(x = Site, y = n, fill = marsh_edge_type)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = n), vjust = -0.4, size = 4) +
  facet_wrap(~ marsh_edge_type, scales = "free_x") +
  scale_fill_manual(values = marsh_edge_colors) +
  labs(
    title = "Sample Size by Site, Ramped vs. Scarped (Averaged)",
    y = "Count of Observations (rows)", x = "Site",
    fill = "Marsh Edge Type"
  ) +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))
rq1_edge_counts_site_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_edge_counts_site_avg.png",
       plot = rq1_edge_counts_site_avg,
       width = 8, height = 5, dpi = 300, limitsize = FALSE)

edge_counts_total_avg <- master_tier5_clean %>%
  count(marsh_edge_type, name = "n")

rq1_edge_counts_total_avg = ggplot(edge_counts_total_avg,
                                aes(x = marsh_edge_type, y = n, fill = marsh_edge_type)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = n), vjust = -0.4, size = 5) +
  scale_fill_manual(values = marsh_edge_colors) +
  labs(
    title = "Total Sample Size, Ramped vs. Scarped (Averaged)",
    y = "Count of Observations (rows)", x = "Marsh Edge Type"
  ) +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))
rq1_edge_counts_total_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_edge_counts_total_avg.png",
       plot = rq1_edge_counts_total_avg,
       width = 6, height = 5, dpi = 300, limitsize = FALSE)
```

## Distance


```{r, ADDED averaged companions for rq1_dist, rq1_dist_noG, rq1_slope}

rq1_dist_avg = ggplot(master_tier5_clean, aes(x = factor(Dist), y = Flux, color = Tran, fill = Tran)) +
  geom_violin(alpha = 0.3, position = position_dodge(0.8)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
              size = 1.2, alpha = 0.4) +
  facet_wrap(~Site, scales = "free_y", ncol = 2) +
 scale_color_manual(values = site_colors) + 
  theme_minimal(base_size = 20) +
  labs(x = "Distance", y = expression(Flux~(g/m^2/day)),
       color = "Transect", fill = "Transect") +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 18),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face = "bold"),
        plot.title = element_text(size = 20, face = "bold"))
rq1_dist_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq1_dist_avg.png",
       plot = rq1_dist_avg,
       width = 13.33,
       height = 7.5,
       dpi = 300,
       limitsize = FALSE)

rq1_dist_noG_avg = ggplot(filter(master_tier5_clean, Site != "GiantMarsh"), 
                      aes(x = factor(Dist), y = Flux, color = Tran, fill = Tran)) +
  geom_violin(alpha = 0.3, position = position_dodge(0.8)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
              size = 1.2, alpha = 0.4) +
  facet_wrap(~Site, scales = "free_y") +
 scale_color_manual(values = site_colors) + 
  theme_minimal(base_size = 20) +
  labs(x = "Distance", y = expression(Flux~(g/m^2/day)),
       color = "Transect", fill = "Transect") +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 18),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face = "bold"))
rq1_dist_noG_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_dist_noG_avg.png",
       plot = rq1_dist_noG_avg,
       width = 13.33,
       height = 7.5,
       dpi = 300,
       limitsize = FALSE)

rq1_slope_avg = ggplot(master_tier5_clean, 
                      aes(x = factor(slope_fitted_elevation_m), y = Flux, color = Tran, fill = Tran)) +
  geom_violin(alpha = 0.3, position = position_dodge(0.8)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8),
              size = 1.2, alpha = 0.4) +
  facet_wrap(~Site, scales = "free_y") +
 scale_color_manual(values = site_colors) + 
  theme_minimal(base_size = 20) +
  labs(x = "Slope Per Transect (m)", y = expression(Flux~(g/m^2/day)),
       color = "Transect", fill = "Transect") +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold", size = 18),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 16, face = "bold"))
rq1_slope_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_slope_avg.png",
       plot = rq1_slope_avg,
       width = 13.33,
       height = 7.5,
       dpi = 300,
       limitsize = FALSE)
```


```{r, ADDED averaged companion for rq1_ramped/rq1_scarped}
# NOTE: mirroring a pre-existing bug in the byrep version, not fixing
# it here -- rq1_scarped's scale_fill_manual() line above is missing a
# trailing "+", so everything after it (theme_minimal/labs/theme) is
# NOT part of the plot object. rq1_scarped as actually assigned only
# has geom_violin+geom_jitter+scale_fill_manual, no title/theme. Same
# applies to rq1_scarped_avg below, for parity -- flagging rather than
# silently fixing since that's a separate change from what was asked.

master_tier5_clean <- master_tier5_clean %>%
  mutate(Site_Tran = paste(Site, Tran, sep = " - "))

rq1_ramped_avg = ggplot(filter(master_tier5_clean, marsh_edge_type == "Ramped"), 
                    aes(x = factor(Dist), y = Flux, color = Site_Tran, fill = Site_Tran)) +
  geom_violin(alpha = 0.3, position = position_dodge(0.8)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15, dodge.width = 1.2),
              size = 0.8, alpha = 0.3) +
  scale_fill_manual(values = site_colors) + 
  theme_minimal(base_size = 13) +
  labs(x = "Distance", y = expression(Flux~(g/m^2)),
       color = "Site - Transect", fill = "Site - Transect",
       title = "Sediment Deposition Rates Across Ramped Sites (Averaged)") +
  theme(legend.position = "bottom")
rq1_ramped_avg

rq1_scarped_avg = ggplot(filter(master_tier5_clean, marsh_edge_type == "Scarped"), 
                     aes(x = factor(Dist), y = Flux, color = Site_Tran, fill = Site_Tran)) +
  geom_violin(alpha = 0.3, position = position_dodge(0.8)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15, dodge.width = 1.2),
              size = 0.8, alpha = 0.3) +
  scale_fill_manual(values = site_colors)
  theme_minimal(base_size = 13) +
  labs(x = "Distance", y = expression(Flux~(g/m^2)),
       color = "Site - Transect", fill = "Site - Transect",
       title = "Sediment Deposition Rates Across Scarped Sites (Averaged)") +
  theme(legend.position = "bottom")
rq1_scarped_avg

```

## Elevation (z*)


```{r, ADDED averaged companion for rq1_z}

rq1_z_avg = ggplot(master_tier5_clean, aes(x = z_star, y = Flux, color = Tran)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_rug(sides = "b", alpha = 0.3, length = unit(0.03, "npc")) +
  facet_wrap(~Site, scales = "free") +
  scale_color_manual(values = c("T1" = "#D85A30", "T2" = "#378ADD")) +
  theme_minimal(base_size = 16) +
  labs(x = "Relative Elevation (Z*)", y = expression(Flux~(g/m^2/day)),
       color = "Transect") +
  theme(strip.text = element_text(face = "bold", size = 16),
        axis.title = element_text(face = "bold", size = 15),
        axis.text = element_text(size = 12))
rq1_z_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\rq1_z_avg.png",
       plot = rq1_z_avg,
       width = 13.33,
       height = 7.5,
       dpi = 300,
       limitsize = FALSE)


```


```{r, ADDED averaged companions for the z* models}

model_avg <- lmer(log(Flux) ~ z_star * marsh_edge_type + (1|Site) + (1|Loc_ID) + (1|Study.Phase), 
              data = master_tier5_clean)
summary(model_avg)

model_full_avg <- lmer(log(Flux) ~ z_star * marsh_edge_type + 
                     (1 | Site) + (1|Loc_ID) + (1 | Study.Phase),  
                   data = master_tier5_clean,
                   REML = FALSE)
summary(model_full_avg)

r.squaredGLMM(model_full_avg)

emtrends(model_full_avg, ~ marsh_edge_type, var = "z_star")

# FIX: (1|Loc_ID) added, same as model_full_glmm above
model_full_glmm_avg <- glmmTMB(
  Flux ~ z_star * marsh_edge_type + (1| Site) + (1|Loc_ID) + (1|Study.Phase), family = Gamma(link = "log"), 
  data = master_tier5_clean
)

summary(model_full_glmm_avg)

sim_res_model_full_avg <- simulateResiduals(model_full_glmm_avg)
plot(sim_res_model_full_avg)
testDispersion(sim_res_model_full_avg)

emtrends(model_full_glmm_avg, ~ marsh_edge_type, var = "z_star")

```


## Distance models


```{r, ADDED averaged companion for distance model prep}
shapiro.test(avgdist_avg$Flux)
shapiro.test(log(avgdist_avg$Flux))

qqnorm(avgdist_avg$Flux)
qqline(avgdist_avg$Flux)
```


```{r, ADDED averaged companions for the distance models}

distmodel_avg <- glmer(Flux ~ Dist + (1|Site/Tran) + (1|Loc_ID), data = avgdist_avg)
summary(distmodel_avg)

logdistmodel_avg <- lmer(log(Flux) ~ Dist + (1|Site/Tran) + (1|Loc_ID), data = avgdist_avg)
summary(logdistmodel_avg)

plot(logdistmodel_avg)
qqnorm(resid(logdistmodel_avg))
qqline(resid(logdistmodel_avg))

outliers_avg <- which(abs(scale(resid(logdistmodel_avg))) > 2.5)
distmodel_no_outliers_avg <- lmer(log(Flux) ~ Dist + (1|Site/Tran) + (1|Loc_ID),
                              data = avgdist_avg[-outliers_avg, ])
summary(logdistmodel_avg)
summary(distmodel_no_outliers_avg)

distmodelcont_avg <- lmer(log(Flux) ~ Dist + (1|Site/Tran) + (1|Loc_ID), data = avgdist_avg)
summary(distmodelcont_avg)
```


```{r, ADDED averaged companion for dist glmm model}

distmodel_glmm_avg <- glmmTMB(
  Flux ~ Dist + (1|Site/Tran) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = avgdist_avg
)
summary(distmodel_glmm_avg)

sim_res_dist_avg <- simulateResiduals(distmodel_glmm_avg)
plot(sim_res_dist_avg)
testDispersion(sim_res_dist_avg)

```


```{r, ADDED averaged companion for distsitemodel}
distsitemodel_avg <- lmer(log(Flux) ~ Dist + Site + Dist*Site + (1|Site:Tran) + (1|Loc_ID), data = avgdist_avg)
summary(distsitemodel_avg)

avgdist_avg$predicted <- predict(distsitemodel_avg)
preds_avg <- ggpredict(distsitemodel_avg, terms = c("Dist [all]", "Site"), back.transform = FALSE)
```


```{r, ADDED averaged companion for dist site model glmm}

distsitemodel_glmm_avg <- glmmTMB(
  Flux ~ Dist + Site + Dist*Site + (1|Site:Tran) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = avgdist_avg
)
summary(distsitemodel_glmm_avg)

sim_res_distsite_avg <- simulateResiduals(distsitemodel_glmm_avg)
plot(sim_res_distsite_avg)
testDispersion(sim_res_distsite_avg)

```

## Transect x Site


```{r, ADDED averaged companion for transect x site model}
transite_avg <- lmer(log(Flux) ~ Tran + Site + Tran*Site + (1|Dist) + (1|Loc_ID), data = avgdist_avg)
summary(transite_avg)

tran_site_avg = ggplot(avgdist_avg, aes(x = as.factor(Tran), y = log(Flux), fill = Site, colour = Site)) +
    geom_boxplot(alpha = 0.4, outlier.shape = NA) +
    geom_sina(alpha = 0.6, size = 2) +
    labs(x = "Transect", y = "Log of Average Flux (g/cm^2/day)",
         fill = "Site", colour = "Site") +
    scale_color_manual(values = site_colors) + 
    scale_fill_manual(values = site_colors) +
    theme_classic(base_size = 18) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18))

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\tran_site_avg.png", 
       plot = tran_site_avg, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```

```{r, ADDED averaged companion for transect and site glmm}
transite_glmm_avg <- glmmTMB(
  Flux ~ Tran + Site + Tran*Site + (1|Dist) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = avgdist_avg
)
summary(transite_glmm_avg)
```


```{r, ADDED averaged companion for transite1}
transite1_avg <- lmer(log(Flux) ~ Tran + Site + Tran*Site +
                    (1|Dist) + (1|Loc_ID) + (1|Study.Phase),
                  data = master_tier5_clean)
summary(transite1_avg)

ggplot(master_tier5_clean, aes(x = as.factor(Tran), y = log(Flux), fill = Site, colour = Site)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA) +
  geom_sina(alpha = 0.6, size = 2) +
  facet_wrap(~Site, scales = "free_x") +
  labs(x = "Transect", y = "Log of Average Flux (g/cm^2/day)",
       fill = "Site", colour = "Site") +
  scale_color_manual(values = marsh_edge_colors) + 
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(hjust = 1, size = 12))
```


```{r, ADDED averaged companion for transite1_glmm}
transite1_glmm_avg <- glmmTMB(
  Flux ~ Tran + Site + Tran*Site + (1|Dist) + (1|Loc_ID) + (1|Study.Phase),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)
summary(transite1_glmm_avg)
```


## Marsh edge type and orientation


```{r, ADDED averaged companion for marsh edge type and orientation model prep}
shapiro.test(avgtotal_avg$Flux)
shapiro.test(log(avgtotal_avg$Flux))

qqnorm(avgtotal_avg$Flux)
qqline(avgtotal_avg$Flux)

shapiro.test(resid(distsitemodel_avg))
qqnorm(resid(distsitemodel_avg))
qqline(resid(distsitemodel_avg))
```


```{r, ADDED averaged companions for marsh edge type and orientation glmm models}

marsh_edge_model_edge_glmm_avg <- glmmTMB(
  Flux ~ marsh_edge_type + (1|Dist) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)
summary(marsh_edge_model_edge_glmm_avg)

marsh_edge_model_sitetran_glmm_avg <- glmmTMB(
  Flux ~ marsh_edge_type + (1|Site/Tran) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)

marsh_edge_model_orient_dist_glmm_avg <- glmmTMB(
  Flux ~ marsh_orientation + (1|Dist) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)

emmeans(marsh_edge_model_edge_glmm_avg, pairwise ~ marsh_edge_type, adjust = "tukey", type = "response")

```


```{r, ADDED averaged companions for marsh edge type and orientation lmer models}
marsh_edge_model_edge_avg <- lmer(Flux ~ marsh_edge_type + (1|Dist) + (1|Loc_ID), data = master_tier5_clean)
summary(marsh_edge_model_edge_avg)

marsh_edge_model_sitetran_avg <- lmer(Flux ~ marsh_edge_type + (1|Site/Tran) + (1|Loc_ID), data = master_tier5_clean)
summary(marsh_edge_model_sitetran_avg)

marsh_edge_model_orient_dist_avg <- lmer(Flux ~ marsh_orientation + (1|Dist) + (1|Loc_ID), data = master_tier5_clean)
summary(marsh_edge_model_orient_dist_avg)

marsh_edge_model_orient_sitetran_avg <- lmer(Flux ~ marsh_orientation + (1|Site/Tran) + (1|Loc_ID), data = master_tier5_clean)
summary(marsh_edge_model_orient_sitetran_avg)

emmeans(marsh_edge_model_edge_avg, pairwise ~ marsh_edge_type, adjust = "tukey")

```


```{r, ADDED averaged companion for marsh_group_model}
marsh_group_model_avg <- lmer(Flux ~ Marsh_Group + (1 | Site/Tran) + (1|Loc_ID), data = marsh_group_flux_avg)
summary(marsh_group_model_avg)
anova(marsh_group_model_avg)
```


```{r, ADDED averaged companion for marsh_group_model_glmm}
marsh_group_model_glmm_avg <- glmmTMB(
  Flux ~ Marsh_Group + (1 | Site/Tran) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = marsh_group_flux_avg
)
summary(marsh_group_model_glmm_avg)

Anova(marsh_group_model_glmm_avg, type = "II")

```


```{r, ADDED averaged companion for the slope-as-covariate lmer models}
deposition_model_slope_avg <- lmer(Flux ~ z_star + local_slope + (1 | Site) + (1|Loc_ID), data = master_tier5_clean)
summary(deposition_model_slope_avg)

deposition_model_base_avg <- lmer(Flux ~ z_star + (1 | Site) + (1|Loc_ID), data = master_tier5_clean)
anova(deposition_model_base_avg, deposition_model_slope_avg)
```


```{r, ADDED averaged companion for the slope-as-covariate glmm models}

deposition_model_slope_glmm_avg <- glmmTMB(
  Flux ~ z_star + local_slope + (1 | Site) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)

deposition_model_base_glmm_avg <- glmmTMB(
  Flux ~ z_star + (1 | Site) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)

anova(deposition_model_base_glmm_avg, deposition_model_slope_glmm_avg)

```


```{r, ADDED averaged companion for marsh edge type and orientation plots}
ggplot(avgtotal_avg, aes(x = marsh_edge_type, y = log(Flux), fill = marsh_orientation)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA) +
  geom_sina(alpha = 0.6, size = 2, aes(colour = marsh_orientation)) +
  labs(x = "Marsh Edge Type", y = "Log of Average Flux (Log of Average Flux (g/cm^2/day))",
       fill = "Marsh Orientation", colour = "Marsh Orientation") +
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18))

ggplot(master_tier5_clean, aes(x = marsh_edge_type, y = log(Flux), fill = marsh_orientation)) +
  geom_violin(alpha = 0.6) +
  geom_boxplot(width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = c("East.facing" = "#E07070",
                               "South.facing" = "#70B870",
                               "West.facing" = "#7090D0")) +
  theme_bw()

basic_rq1_avg = ggplot(avgtotal_avg, aes(x = marsh_edge_type, y = Flux, fill = marsh_edge_type)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(alpha = 0.4, width = 0.2, size = 0.8) +
  scale_y_log10() +
  labs(title = str_wrap("Sediment Flux by Marsh Edge Type (Averaged)", width = 40),
       y = "Average Flux (g/m^2/day)", x = "Marsh Edge Type") +
  theme_bw() +
  scale_fill_manual(values = marsh_edge_colors) + 
  theme(legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 13),
        plot.title = element_text(hjust = 0.5))
basic_rq1_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\basic_rq1_avg.png", 
       plot = basic_rq1_avg, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

basic_rq1_tran_avg = ggplot(avgdist_avg, aes(x = Tran, y = Flux, fill = Site)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(alpha = 0.4, width = 0.2, size = 0.8) +
  scale_y_log10() +
  labs(title = str_wrap("Sediment Flux by Marsh Edge Type (Averaged)", width = 40),
       y = "Average Flux (g/m^2/day)", x = "Transect") +
  theme_bw() +
  scale_fill_manual(values = site_colors) + 
  theme(legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 13),
        plot.title = element_text(hjust = 0.5))
basic_rq1_tran_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\basic_rq1_tran_avg.png", 
       plot = basic_rq1_tran_avg, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

basic_rq1_tran_noG2_avg = ggplot(avgdist_avg %>% filter(!(Site == "GiantMarsh" & Tran == "T2")), aes(x = Tran, y = Flux, fill = Site)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(alpha = 0.4, width = 0.2, size = 0.8) +
  scale_y_log10() +
  labs(title = str_wrap("Sediment Flux by Marsh Edge Type (Averaged)", width = 40),
       y = "Average Flux (g/m^2/day)", x = "Transect") +
  theme_bw() +
  scale_fill_manual(values = site_colors) + 
  theme(legend.position = "right",
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 13),
        plot.title = element_text(hjust = 0.5))
basic_rq1_tran_noG2_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\basic_rq1_tran_noG2_avg.png", 
       plot = basic_rq1_tran_noG2_avg, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

ggplot(avgtotal_avg, aes(x = marsh_orientation, y = Flux, fill = marsh_orientation)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_jitter(alpha = 0.4, width = 0.2, size = 0.8) +
  scale_y_log10() +
  labs(title = str_wrap("Sediment Flux by Marsh Orientation (Averaged)", width = 40),
       y = "Log of Average Flux (g/cm^2/day)", x = "Marsh Orientation") +
  theme_bw() +
  scale_color_manual(values = marsh_edge_colors) + 
  theme(legend.position = "none",
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 13),
        plot.title = element_text(hjust = 0.5))
```


```{r, ADDED averaged companion for p_with_48/p_without_48}

p_with_48_avg <- ggplot(master_tier5_with48, aes(x = Dist, y = Flux)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  labs(title = "Flux vs. Distance — With 48m Point (Averaged)", x = "Distance (m)", y = "Flux") +
  theme_bw()

p_without_48_avg <- ggplot(filter(master_tier5_clean, Dist != 48), aes(x = Dist, y = Flux)) +
  geom_point(alpha = 0.3, color = "darkred") +
  labs(title = "Flux vs. Distance — Without 48m Point (Averaged)", x = "Distance (m)", y = "Flux") +
  theme_bw()

p_with_48_avg
p_without_48_avg
```


```{r, ADDED averaged companion for plot_zstar/plot_slope}

plot_zstar_avg <- ggpredict(deposition_model_slope_avg, terms = "z_star") %>%
  plot() +
  labs(title = "Predicted Flux by Z-star (Averaged)", x = "Z-star", y = "Predicted Flux") +
  theme_bw()

plot_slope_avg <- ggpredict(deposition_model_slope_avg, terms = "local_slope") %>%
  plot() +
  labs(title = "Predicted Flux by Local Slope (Averaged)", x = "Local Slope", y = "Predicted Flux") +
  theme_bw()

plot_zstar_avg
plot_slope_avg

```


```{r, ADDED averaged companion for the marsh_group emmeans plot}
summary(marsh_group_model_avg)

emm_marsh_group_avg <- emmeans(marsh_group_model_avg, ~ Marsh_Group)
emm_marsh_group_avg_df <- as.data.frame(emm_marsh_group_avg)

ggplot(emm_marsh_group_avg_df, aes(x = Marsh_Group, y = emmean)) +
  geom_col(fill = "steelblue", width = 0.5) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.15) +
  labs(
    title = "Estimated Marginal Means of Flux by Geomorphic Grouping (Averaged)",
    y = "Estimated Flux", x = "Marsh Group"
  ) +
  theme_bw()
```


```{r, ADDED averaged companion for BCtransite}
BCtransite_avg <- lmer(log(Flux) ~ Tran + Site + Tran*Site + (1|Dist) + (1|Loc_ID), data = avgdist_protected_avg)
summary(BCtransite_avg)

BCtransite_plot_avg = ggplot(avgdist_protected_avg, aes(x = as.factor(Tran), y = log(Flux), fill = Site, colour = Site)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA) +
  geom_sina(alpha = 0.6, size = 2) +
  labs(x = "Transect", y = "Log of Average Flux (g/cm^2/day)",
       fill = "Site", colour = "Site") +
  scale_color_manual(values = site_colors) +
  scale_fill_manual(values = site_colors) + 
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18))
BCtransite_plot_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\BCtransite_avg.png", 
       plot = BCtransite_plot_avg,
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```


```{r, ADDED averaged companion for p_flux_zstar}

p_flux_zstar_avg <- ggplot(master_tier5_clean, aes(x = z_star, y = Flux, color = Site)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_manual(values = site_colors) +
   scale_y_log10() +
  labs(
    title = str_wrap("Daily Sediment Flux vs. Elevation (z*) Across 2 Ramped and 2 Scarped Marshes, 2025-2026 (Averaged)", width = 40),
    x = "z*",
    y = expression(Flux~(g/m^2/day)),
    color = "Site"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
p_flux_zstar_avg


ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\p_flux_zstar_avg.png", 
       plot = p_flux_zstar_avg,
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```


```{r, ADDED averaged companion for p_flux_dist}

p_flux_dist_avg <- ggplot(master_tier5_clean, aes(x = Dist, y = Flux, color = Site)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_manual(values = site_colors) +
   scale_y_log10() +
  labs(
    title = str_wrap("Daily Sediment Flux vs. Distance Across 2 Ramped and 2 Scarped Marshes, 2025-2026 (Averaged)", width = 40),
    x = "Distance (m)",
    y = expression(Flux~(g/m^2/day)),
    color = "Site"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
p_flux_dist_avg


ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\p_flux_dist_avg.png", 
       plot = p_flux_dist_avg,
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```


```{r, ADDED averaged companion for p_flux_dist_shape}

master_tier5_clean$Site <- factor(master_tier5_clean$Site, 
                   levels = c("GiantMarsh", "CorteMadera", "SanPablo", "Buck'sLanding"))

p_flux_dist_shape_avg <- master_tier5_clean %>%
  mutate(Dist_factor = factor(Dist, levels = c(0.5, 2, 6, 12, 18, 24))) %>%
  ggplot(aes(x = Dist_factor, y = Flux, fill = Site, color = Site)) +
  geom_violin(alpha = 0.4, position = position_dodge(width = 0.8), scale = "width") +
  geom_jitter(size = 1, alpha = 0.3, position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.8)) +
  scale_y_log10() +
  scale_fill_manual(values = site_colors) +
  scale_color_manual(values = site_colors) +
  facet_wrap(~ Site, ncol = 1) +
  labs(
    title = str_wrap("Distribution Shape of Sediment Flux by Distance, Ramped vs. Scarped (Averaged)", width = 40),
    x = "Distance (m)",
    y = expression(Flux~(g/m^2/day)),
    fill = "Site", color = "Site"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
p_flux_dist_shape_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\p_flux_dist_shape_avg.png", 
       plot = p_flux_dist_shape_avg,
       width = 10, 
       height = 14, 
       dpi = 300,
       limitsize = FALSE)
```


```{r, ADDED averaged companions for the three final distance/edge-type models}

m_glmm_avg <- glmmTMB(Flux ~ Dist * marsh_edge_type + (1 | Site),
                   family = Gamma(link = "log"),
                   data = master_tier5_clean)

summary(m_glmm_avg)

m_no_giant_avg <- glmmTMB(Flux ~ factor(Dist) * marsh_edge_type + (1 | Site),
                       family = Gamma(link = "log"),
                       data = master_tier5_clean %>% filter(Site != "GiantMarsh"))

summary(m_no_giant_avg)

master_close_dist_avg <- master_tier5_clean %>%
  filter(Dist %in% c(0.5, 2))

m_close_avg <- glmmTMB(Flux ~ Dist * marsh_edge_type + (1 | Site),
                    family = Gamma(link = "log"),
                    data = master_close_dist_avg)
summary(m_close_avg)
```


```{r, ADDED averaged companion for p_edge_zone_raincloud}

p_edge_zone_raincloud_avg <- master_tier5_clean %>%
  filter(Dist %in% c(0.5, 2)) %>%
  mutate(Dist_factor = factor(Dist, levels = c(0.5, 2), labels = c("0.5 m", "2 m"))) %>%
  ggplot(aes(x = Site, y = Flux, fill = Site, color = Site)) +
  stat_halfeye(adjust = 0.5, width = 0.6, justification = -0.2, alpha = 0.5, .width = 0, point_colour = NA) +
  geom_boxplot(width = 0.12, outlier.shape = NA, alpha = 0.6) +
  scale_y_log10() +
  scale_fill_manual(values = site_colors) +
  scale_color_manual(values = site_colors) +
  facet_wrap(~ Dist_factor) +
  labs(
    title = str_wrap("Sediment Flux Distribution Shape at the Marsh Edge (0.5m & 2m) (Averaged)", width = 40),
    x = "Site", y = expression(Flux~(g/m^2/day))
  ) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))
p_edge_zone_raincloud_avg

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\p_edge_zone_raincloud_avg.png", 
       plot = p_edge_zone_raincloud_avg,
       width = 9, 
       height = 6, 
       dpi = 300,
       limitsize = FALSE)
```


```{r, ADDED averaged companion for site_stats}
site_stats_avg <- make_flux_table(
  master_tier5_clean, 
  group_vars = "Site", 
  title = "Descriptive Statistics of Sediment Flux by Site (Averaged)",
  subtitle = "Flux measured in g/m²/day"
)
```


```{r, ADDED averaged companion for the site_stats image export}
ft_avg <- site_stats_avg$data %>%
  flextable() %>%
  set_caption("Descriptive Statistics of Sediment Flux by Site (Averaged)") %>%
  colformat_double(digits = 3) %>%
  autofit()

save_as_image(ft_avg, path = "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\site_flux_stats_avg.png")
```


```{r, ADDED averaged companion for the dispersion model}

distsitemodel_glmm_dispersion_avg <- glmmTMB(
  Flux ~ Dist * Site + (1|Site:Tran) + (1|Loc_ID),
  dispformula = ~ Site,
  family = Gamma(link = "log"),
  data = avgdist_avg
)
summary(distsitemodel_glmm_dispersion_avg)

AIC(distsitemodel_glmm_avg, distsitemodel_glmm_dispersion_avg)

```


```{r, ADDED averaged companion for the GiantMarsh sensitivity check}

avgdist_noGM_avg <- avgdist_avg %>% filter(Site != "GiantMarsh")

distsitemodel_glmm_noGM_avg <- glmmTMB(
  Flux ~ Dist * Site + (1|Site:Tran) + (1|Loc_ID),
  family = Gamma(link = "log"),
  data = avgdist_noGM_avg
)

summary(distsitemodel_glmm_avg)      # averaged, all 4 sites
summary(distsitemodel_glmm_noGM_avg) # averaged, GiantMarsh excluded
```

## Output tables (ADDED)

```{r, per-site distance stats tables}
# Splits the Site x Dist descriptive table (tier5_dist_site_stats) into one
# table per site. Flux > 0, 48 m removed at load (so, unlike step6's
# combined tier5_dist_site_stats table, there's no 48 m row).
# Output: dist_stats_<Site>_avg.png

rq1_img_dir <- "C:/Users/savan/OneDrive - San Francisco State University/ProjectData/GitHub/DepositionSFB/images/rq_1"

dist_site_tables_avg <- save_site_split_tables(
  master_tier5_clean, group_var = "Dist", group_label = "Distance (m)",
  file_stub = "dist_stats", tier = "avg", out_dir = rq1_img_dir
)
```

```{r, save fitted models}
# The GLMM output-table images need BOTH tiers of each model, which now
# live in separate files. So each tier file saves its fitted models here,
# and RQ1-spatial_tables.Rmd reads both and makes the images -- no
# refitting, so that file knits in seconds. Knit this file and
# RQ1-spatial_byrep.Rmd before RQ1-spatial_tables.Rmd. Any model that
# wasn't fit is saved as NULL and skipped by the tables file.

rq1_spatial_model_names_avg <- c(
  "model_full_glmm_avg",
  "distmodel_glmm_avg",
  "distsitemodel_glmm_avg",
  "distsitemodel_glmm_dispersion_avg",
  "distsitemodel_glmm_noGM_avg",
  "transite_glmm_avg",
  "transite1_glmm_avg",
  "marsh_edge_model_edge_glmm_avg",
  "marsh_edge_model_sitetran_glmm_avg",
  "marsh_edge_model_orient_dist_glmm_avg",
  "marsh_group_model_glmm_avg",
  "deposition_model_base_glmm_avg",
  "deposition_model_slope_glmm_avg",
  "m_glmm_avg",
  "m_no_giant_avg",
  "m_close_avg"
)
rq1_spatial_models_avg <- mget(rq1_spatial_model_names_avg, ifnotfound = list(NULL))

saveRDS(rq1_spatial_models_avg,
        file.path(knitr::opts_knit$get("root.dir"), "rq1_spatial_models_avg.rds"))

sapply(rq1_spatial_models_avg, Negate(is.null))   # FALSE = not found, will be skipped
```

---

**Note on scope:** every model-fitting call in this file now has an `_avg` companion, using the same formula and consolidated averaged-side data objects (`master_tier5_clean`, `avgdist_avg`, `avgtotal_avg`, `avgdist_protected_avg`, `marsh_group_flux_avg`, `avgdist_noGM_avg`). Model-integral diagnostics (summary, Anova, emmeans, DHARMa residual checks, nested-model anova comparisons) were duplicated alongside each model.

**Plot companions (updated in the byrep/avg check):** every saved figure in this file now has an `_avg` companion, and every byrep output file ends in `_byrep` (byrep plot titles are unchanged; `_avg` plot titles carry "(Averaged)"). `rq1_boxplot`/`rq1_boxplot_bysite` were the last two without one and now have an averaged chunk. The only plots still without a saved `_avg` file are the ones that aren't saved as files at all on the byrep side either (the unassigned ggpredict z* plot and a few quick-look plots printed to the viewer).

