---
title: "RQ1 - Temporal (Claude)"
format: pdf
editor: visual
---

```{=html}
<!--
Companion to RQ1-spatial.Rmd, split out this session. This file holds
everything from master_thesis_regressions.Rmd and master_thesis_visualizations.qmd
that looks at Flux across Study.Phase, Month, date, or Hydro_Period,
rather than across site/edge-type/distance/elevation.

Moved in from master_thesis_regressions.Rmd:
- "derive Month for reference models" (Month factor derivation)
- "rq1 time" chunk -> rq1_time_trend, rq1_time (Flux by Month)
- two of the three plots from "transect x site model with replicate and
  phase" -- Study.Phase sina plot and Dist x Study.Phase facet plot.
  (transite1 model itself and its Tran-by-Site plot stayed in
  RQ1-spatial.Rmd, since the model's actual question is Tran x Site.)

Everything else below was already rebuilt in master_thesis_visualizations.qmd
this session (bug fixes noted there carried over unchanged): the
Study.Phase trend-by-site/transect/distance-bin plots and the phenological
period section.

Renamed Hydro_Period -> Phen_Period throughout: that column no longer
exists under the old name, and rather than keep reconstructing it from
Study.Phase (a guess), this now uses the real Phen_Period column already
on master_tier5, with the same Late_Dry/Tran/Wet/Growth levels/order as
the old Hydro_Period (confirmed).
-->
```

```{r, load packages and working directory}

source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\setup.R")
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\plotting.R")

# Shared fixed/random-effects rule, family decision, and naming
# convention -- see that file's header before editing any model below.
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\modeling_conventions (Claude).R")

# Definitions for every statistical function used in this file -- see
# the "Statistical Tools Used in This File" note below, or
# statistical_tools_glossary (Claude).R for the fuller version.
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\statistical_tools_glossary (Claude).R")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/savan/OneDrive - San Francisco State University/ProjectData/GitHub/DepositionSFB/data')

```

```{=html}
<!--
STATISTICAL TOOLS USED IN THIS FILE
(full definitions in statistical_tools_glossary (Claude).R)

- lmer() [lme4] -- Gaussian mixed model; SENSITIVITY check version of
  edge_time_model_glmm (Rule 2).
- glmmTMB() [glmmTMB] -- PRIMARY model fitter, Gamma(link = "log").
  Includes the ar1() correlation structure in edge_time_model_glmm_ar1,
  for the repeated-monthly-sampling autocorrelation check below.
- Anova() [car] -- Type II significance test on fixed effects.
- anova() [base/lme4] -- fixed-effects table for a single lmer model.
- AIC() -- compares edge_time_model_glmm against its AR1 version
  (non-nested models, so a likelihood-ratio anova() isn't appropriate).
- emmeans() [emmeans] -- model-predicted means by marsh_edge_type x
  Month, with type = "response" to back-transform off the log link.
- simulateResiduals() / testDispersion() [DHARMa] -- simulation-based
  residual diagnostics appropriate for a Gamma GLMM.
- testTemporalAutocorrelation() / recalculateResiduals() [DHARMa] --
  checks whether residuals at consecutive Study.Phase time points are
  correlated, run separately per Loc_ID (site x transect x distance) via
  recalculateResiduals() -- the actual justification for trying the
  ar1() structure, rather than assuming autocorrelation is a problem.
-->
```

```{r, load and prep data}

master_tier5 <- readr::read_csv(
  file.path(knitr::opts_knit$get("root.dir"), "master_tier5.csv")
)

master_tier5_clean <- master_tier5 %>%
    filter(Flux > 0, 
         Dist != 48) %>%
  mutate(
    Study.Phase = factor(Study.Phase, levels = paste0("M", 1:12)),
    Site = factor(Site, levels = c("GiantMarsh", "SanPablo", "Buck'sLanding", "CorteMadera")),
    Tran = factor(Tran),
    marsh_edge_type = factor(marsh_edge_type, levels = c("Scarped", "Ramped")),
    marsh_orientation = as.factor(marsh_orientation), 
    Month = factor(
      Month, levels = c("Jul", "Aug", "Sep", "Oct", "Nov", "Dec",
                                       "Jan", "Feb", "Mar", "Apr", "May", "Jun")), 
    Phen_Period = factor(
      Phen_Period,
      levels = c("Late_Dry\n(Months 1-3)", "Tran\n(Months 4-6)",
                 "Wet\n(Months 7-9)", "Growth\n(Months 10-12)")
    ), 
    Cal_Season = factor(Cal_Season, levels = c("Winter", "Summer")),   # NEW: Lacy et al. (2019) Nov-Feb/Mar-Oct
      Loc_ID = interaction(Site, Tran, Dist, drop = TRUE)   # NEW: stable per-location ID, reused for AR1 grouping and diagnostics
  
)



```

## Flux over calendar date

```{r, rq1 time}

summary_df <- master_tier5_clean %>%
  group_by(Site, Tran, Month) %>%
  summarise(median_flux = median(Flux, na.rm = TRUE),
            q25 = quantile(Flux, 0.25, na.rm = TRUE),
            q75 = quantile(Flux, 0.75, na.rm = TRUE),
            .groups = "drop")

rq1_time_trend = ggplot(summary_df, aes(x = Month, y = median_flux, color = Tran, fill = Tran)) +
  geom_ribbon(aes(ymin = q25, ymax = q75), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  facet_wrap(~Site, scales = "free_y") +
  scale_color_manual(values = transect_colors) +
  scale_fill_manual(values = transect_colors) +
  theme_minimal(base_size = 15) +
  labs(x = "Date Removed from Field", y = expression(Median~Flux~(g/m^2/day)),
       color = "Transect", fill = "Transect") +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold"), 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 18))
rq1_time_trend

ggsave(file.path(knitr::opts_knit$get("root.dir"), "rq1_time_trend.png"),
       plot = rq1_time_trend,
       width = 13.33,
       height = 7.5,
       dpi = 300,
       limitsize = FALSE)

rq1_time = ggplot(master_tier5_clean, aes(x = Month, y = Flux, color = Tran, fill = Tran)) +
  geom_boxplot(aes(group = interaction(Month, Tran)), 
               alpha = 0.3, outlier.shape = NA, 
               position = position_dodge(width = 6)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 1.5, dodge.width = 6),
              size = 0.8, alpha = 0.3) +
  facet_wrap(~Site, scales = "free_y") +
  scale_color_manual(values = c(values = transect_colors)) +
  scale_fill_manual(values = c(values = transect_colors)) +
  theme_minimal(base_size = 15) +
  labs(x = "Date Removed from Field", y = expression(Flux~(g/m^2/day)),
       color = "Transect", fill = "Transect") +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1))
rq1_time

ggsave(file.path(knitr::opts_knit$get("root.dir"), "rq1_time.png"),
       plot = rq1_time,
       width = 13.33,
       height = 7.5,
       dpi = 300,
       limitsize = FALSE)

```

## Flux over Study.Phase

```{r, study phase raw distribution by site}
# Moved from RQ1-spatial.Rmd's transite1 chunk -- shows the raw Flux
# distribution across phase, complementing the summary-line versions below.

log_month = ggplot(master_tier5_clean, aes(x = as.factor(Month), y = log(Flux), color = Site)) +
  geom_point(alpha = 0.4) +
  geom_sina(alpha = 0.6, size = 2) +
  labs(x = "Month of Study", y = "Log Flux (g/cm^2/day)",
       colour = "Site") +
  scale_color_manual(values = c(site_colors)) + 
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18))
log_month 

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\log_month .png", 
       plot = log_month, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

wrapmonth_dist = ggplot(master_tier5_clean, aes(x = Dist, y = log(Flux), color = Site)) +
  facet_wrap("Month") +
  geom_point(alpha = 0.4) +
  geom_sina(alpha = 0.6, size = 2) +
  labs(x = "Distance", y = "Log Flux (g/cm^2/day)",
       colour = "Site") +
  scale_color_manual(values = c(site_colors)) +
  theme_classic(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18)) 
wrapmonth_dist

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\wrapmonth_dist.png", 
       plot = wrapmonth_dist, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```

```{r, aggregate per Study.Phase-Site}

total_summary <- master_tier5_clean %>%
  group_by(Site, Month) %>%
  summarise(
    mean_flux = mean(Flux, na.rm = TRUE),
    se_flux = sd(Flux, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

mean_month_site = ggplot(total_summary, aes(x = Month, y = mean_flux, group = Site, color = Site)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_wrap(~ Site, scales = "free_y") +
  scale_color_manual(values = c(site_colors)) + 
  theme_minimal() +
  labs(x = "Majority Month Deployed", y = expression(Mean~Flux~(g/m^2/day))) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
mean_month_site

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\mean_month_site.png", 
       plot = mean_month_site, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```

```{r, Study Phase by marsh group and transect}

phase_compare <- master_tier5_clean %>%
  mutate(
    Marsh_Group = case_when(
      Site %in% c("SanPablo", "GiantMarsh") ~ "San Pablo + Giant Marsh",
      Site %in% c("Buck'sLanding", "CorteMadera") ~ "Buck's Landing + Corte Madera",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(Marsh_Group)) %>%
  group_by(Marsh_Group, Site, Tran, Month) %>%
  summarise(
    mean_flux = mean(Flux, na.rm = TRUE),
    se_flux = sd(Flux, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

phase_compare_grouped <- phase_compare %>%
  group_by(Marsh_Group, Month) %>%
  summarise(
    mean_flux = weighted.mean(mean_flux, n, na.rm = TRUE),
    se_flux = sqrt(sum(se_flux^2 * n^2, na.rm = TRUE)) / sum(n),
    .groups = "drop"
  )

ggplot(phase_compare_grouped, aes(x = Month, y = mean_flux, color = Marsh_Group, group = Marsh_Group)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  theme_minimal() +
  labs(x = "Majority Month Deployed", y = expression(Mean~Flux~(g/m^2/day)), color = "Marsh Group")

mean_month = ggplot(phase_compare, aes(x = Month, y = mean_flux, color = Tran, group = Tran)) +
  geom_point(size = 3) +
  scale_color_manual(values = c(transect_colors)) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_wrap(~ Site, scales = "free_y") +
  theme_minimal() +
  labs(x = "Month Majority Deployment", y = expression(Mean~Flux~(g/m^2/day)), color = "Transect") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
mean_month

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\mean_month.png", 
       plot = mean_month, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)

```

```{r, Study Phase by distance}

transect_summary2 <- master_tier5_clean %>%
  mutate(Flux_zeroed = ifelse(Flux < 0, 0, Flux)) %>%
  filter(Dist %in% c(0.5, 2, 6, 12, 18, 24)) %>%
  mutate(Dist = factor(Dist, levels = c(0.5, 2, 6, 12, 18, 24))) %>%
  group_by(Dist, Site, Month) %>%
  summarise(
    mean_flux = mean(Flux_zeroed, na.rm = TRUE),
    se_flux = sd(Flux_zeroed, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

dist_month <- ggplot(transect_summary2, aes(x = Month, y = mean_flux, color = Site, group = Site)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = pmax(mean_flux - se_flux, 0.001), ymax = mean_flux + se_flux),
                width = 0.2, linewidth = 0.5) +
  facet_wrap(~ Dist, scales = "free_y", ncol = 2) +
  scale_y_log10() +
  scale_color_manual(values = c(site_colors)) + 
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 13, face = "bold"),
        legend.position = "bottom") +
  labs(
    title = str_wrap("Mean Sediment Flux Across Study Phase by Distance from Marsh Edge", width = 40),
    subtitle = "Negative flux values set to zero",
    x = "Majority Month Deployed", y = expression(Mean~Flux~(g/m^2/day)), color = "Site"
  )
dist_month

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\dist_month.png",
       plot = dist_month,
       width = 16,
       height = 14,
       dpi = 300,
       limitsize = FALSE)

dist_plot_mean <- ggplot(transect_summary2, aes(x = Dist, y = mean_flux, color = Month, group = Site)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = pmax(mean_flux - se_flux, 0.001), ymax = mean_flux + se_flux),
                width = 0.2, linewidth = 0.5) +
  facet_wrap(~ Site, scales = "free_y", ncol = 2) +
  scale_y_log10() +
  scale_color_manual(values = c(site_colors)) + 
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 13, face = "bold"),
        legend.position = "bottom") +
  labs(
    title = str_wrap("Mean Sediment Flux Across Study Phase by Distance from Marsh Edge", width = 40),
    subtitle = "Negative flux values set to zero",
    x = "Distance from Marsh Edge (m)", y = expression(Mean~Flux~(g/m^2/day)), color = "Site"
  )
dist_plot_mean

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\dist_plot_mean.png",
       plot = dist_plot_mean,
       width = 16,
       height = 14,
       dpi = 300,
       limitsize = FALSE)


dist_plot_simple <- ggplot(transect_summary2, aes(x = Dist, y = mean_flux, color = Site)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = pmax(mean_flux - se_flux, 0.001), ymax = mean_flux + se_flux),
                width = 0.2, linewidth = 0.5) +
  facet_wrap(~ Site, scales = "free_y", ncol = 2) +
  scale_y_log10() +
  scale_color_manual(values = c(site_colors)) + 
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 13, face = "bold"),
        legend.position = "bottom") +
  labs(
    title = str_wrap("Mean Sediment Flux Across Study Phase by Distance from Marsh Edge", width = 40),
    subtitle = "Negative flux values set to zero",
    x = "Distance from Marsh Edge (m)", y = expression(Mean~Flux~(g/m^2/day)), color = "Site"
  )
dist_plot_simple

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\dist_plot_simple.png",
       plot = dist_plot_simple,
       width = 16,
       height = 14,
       dpi = 300,
       limitsize = FALSE)

transect_summary_no_giant <- master_tier5_clean %>%
  mutate(Flux_zeroed = ifelse(Flux < 0, 0, Flux)) %>%
  filter(Site != "GiantMarsh", Dist %in% c(0.5, 2, 6, 12, 18, 24)) %>%
  mutate(Dist = factor(Dist, levels = c(0.5, 2, 6, 12, 18, 24))) %>%
  group_by(Dist, Site, Month) %>%
  summarise(
    mean_flux = mean(Flux_zeroed, na.rm = TRUE),
    se_flux = sd(Flux_zeroed, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

dist_plot_no_giant <- ggplot(transect_summary_no_giant, aes(x = Month, y = mean_flux, color = Site, group = Site)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2, linewidth = 0.5) +
  facet_wrap(~ Dist, scales = "free_y", ncol = 2) +
  scale_color_manual(values = c(site_colors)) + 
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 13, face = "bold"),
        legend.position = "bottom") +
  labs(
    title = str_wrap("Mean Sediment Flux Across Study Phase by Distance from Marsh Edge", width = 40),
    subtitle = "GiantMarsh removed; negative flux values set to zero",
    x = "Majority Month Deployed", y = expression(Mean~Flux~(g/m^2/day)), color = "Site"
  )
dist_plot_no_giant

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\dist_plot_no_giant.png",
       plot = dist_plot_no_giant,
       width = 16,
       height = 14,
       dpi = 300,
       limitsize = FALSE)

```

## Phenological period

```{r, phen period plots}

phen_summary <- master_tier5_clean %>%
  mutate(Flux_zeroed = ifelse(Flux < 0, 0.001, Flux)) %>%
  group_by(Phen_Period, Site) %>%
  summarise(
    mean_flux = mean(Flux_zeroed, na.rm = TRUE),
    se_flux = sd(Flux_zeroed, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

ggplot(phen_summary, aes(x = Phen_Period, y = mean_flux, color = Site, group = Site)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = pmax(mean_flux - se_flux, 0.001), ymax = mean_flux + se_flux), width = 0.2) +
  scale_y_log10() +
  theme_minimal(base_size = 14) +
  labs(title = "Mean Flux by Phenological Period", subtitle = "Log-scaled y-axis",
       x = "Phenological Period", y = expression(Mean~Flux~(g/m^2/day)), color = "Site")

phen_summary_no_giant <- master_tier5_clean %>%
  filter(Site != "GiantMarsh") %>%
  mutate(Flux_zeroed = ifelse(Flux < 0, 0, Flux)) %>%
  group_by(Phen_Period, Site) %>%
  summarise(
    mean_flux = mean(Flux_zeroed, na.rm = TRUE),
    se_flux = sd(Flux_zeroed, na.rm = TRUE) / sqrt(n()),
    n = n(),
    .groups = "drop"
  )

ggplot(phen_summary_no_giant, aes(x = Phen_Period, y = mean_flux, color = Site, group = Site)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  theme_minimal(base_size = 14) +
  labs(title = "Mean Flux by Phenological Period", subtitle = "GiantMarsh removed",
       x = "Phenological Period", y = expression(Mean~Flux~(g/m^2/day)), color = "Site")

ggplot(phen_summary, aes(x = Phen_Period, y = mean_flux, group = 1)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_flux - se_flux, ymax = mean_flux + se_flux), width = 0.2) +
  facet_wrap(~ Site, scales = "free_y") +
  theme_minimal(base_size = 14) +
  labs(title = "Mean Flux by Phenological Period", x = "Phenological Period", y = expression(Mean~Flux~(g/m^2/day)))
```

```{r, temporal marsh models}
# SENSITIVITY check (Rule 2) -- edge_time_model_glmm below is the
# reported/primary model for this comparison. (1|Site) alone (no Tran)
# is intentional here, not the nesting bug flagged elsewhere: this
# model isn't using Tran as a random effect at all, so there's nothing
# to nest.

edge_time_model <- lmer(
  Flux ~ marsh_edge_type * Month + (1|Site),
  data = master_tier5_clean
)
summary(edge_time_model)
anova(edge_time_model)

```

```{r, edge time glmm}
# PRIMARY model for the edge-type x Month comparison (Rule 2: Gamma GLMM
# on raw Flux).

edge_time_model_glmm <- glmmTMB(
  Flux ~ marsh_edge_type * Month + (1|Site),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)
summary(edge_time_model_glmm)
Anova(edge_time_model_glmm, type = "II")

sim_res_edge_time <- simulateResiduals(edge_time_model_glmm)

locs <- unique(master_tier5_clean$Loc_ID)
for (loc in locs) {
  idx <- which(master_tier5_clean$Loc_ID == loc)
  if (length(idx) < 3) next
  sim_res_loc <- recalculateResiduals(sim_res_edge_time, sel = idx)
  cat("\n---", as.character(loc), "---\n")
  print(testTemporalAutocorrelation(sim_res_loc, time = master_tier5_clean$Study.Phase[idx]))
}


# AR1 version, since this is repeated monthly sampling
master_tier5_clean_ar <- master_tier5_clean %>%
  mutate(Loc_ID = interaction(Site, Tran, Dist, drop = TRUE))

edge_time_model_glmm_ar1 <- glmmTMB(
  Flux ~ marsh_edge_type * Month + ar1(Study.Phase + 0 | Loc_ID), #because month was messy for the linear model-changing to study phase to keep track for categorical sorting
  family = Gamma(link = "log"),
  data = master_tier5_clean_ar
)
summary(edge_time_model_glmm_ar1)

AIC(edge_time_model_glmm, edge_time_model_glmm_ar1)

```

```{r}
library(emmeans)
library(ggplot2)

# FIX: this previously pulled emmeans from edge_time_model (the
# Gaussian SENSITIVITY check), not edge_time_model_glmm (the PRIMARY
# Gamma model per Rule 2) -- meaning the plot actually reported wasn't
# built from the model treated as primary everywhere else in this
# chunk. Switched to the glmm, with type = "response" so emmean/CL come
# back on the Flux scale rather than the log-link scale.

emm_edge_time <- emmeans(edge_time_model_glmm, ~ marsh_edge_type * Month, type = "response")
emm_edge_time_df <- as.data.frame(emm_edge_time)

# emmeans(type = "response") on a Gamma GLMM names the back-transformed
# estimate "response" (not "emmean") and its CI "asymp.LCL"/"asymp.UCL"
# (not "lower.CL"/"upper.CL") -- renamed here so the ggplot code below,
# written for the lmer-style column names, keeps working unchanged.
emm_edge_time_df <- emm_edge_time_df %>%
  rename(emmean = response, lower.CL = asymp.LCL, upper.CL = asymp.UCL)

emm = ggplot(emm_edge_time_df, aes(x = Month, y = emmean, color = marsh_edge_type, group = marsh_edge_type)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(0.3)) +
  labs(
    title = str_wrap("Estimated Marginal Means of Flux by Marsh Edge Type Across Month within 2025-2026", width = 40),
    y = "Estimated Flux", x = "Majority Month Deployed", color = "Edge Type"
  ) +
  theme_bw() +
  scale_color_manual(values = marsh_edge_colors) + 
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
emm 

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\emm.png", 
       plot = emm,
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)


```

```{r}
# SENSITIVITY check: does excluding GiantMarsh change the conclusion?
# Fit only as the Gaussian lmer() here, not re-fit as a Gamma GLMM --
# that's an intentional scope choice (this is a robustness check on the
# primary edge_time_model_glmm result, not a second primary model), but
# flagging it explicitly so it's clear this asymmetry is a decision and
# not an oversight.

master_tier5_clean_noGM <- master_tier5_clean %>%
  filter(Site != "GiantMarsh")

edge_time_model_noGM <- lmer(
  Flux ~ marsh_edge_type * Month + (1|Site),
  data = master_tier5_clean_noGM
)

summary(edge_time_model_noGM)
anova(edge_time_model_noGM)

emm_edge_time_noGM <- emmeans(edge_time_model_noGM, ~ marsh_edge_type * Month)
emm_edge_time_noGM_df <- as.data.frame(emm_edge_time_noGM)

emm_noGM = ggplot(emm_edge_time_noGM_df, aes(x = Month, y = emmean, color = marsh_edge_type, group = marsh_edge_type)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(0.3)) +
  labs(
    title = stringr::str_wrap("Estimated Marginal Means of Flux by Marsh Edge Type Across Study Phase (Excluding Giant Marsh)"), width = 50,
    y = "Estimated Flux", x = "Majority Month Deployed", color = "Edge Type"
  ) +
  theme_bw() +
  scale_color_manual(values = marsh_edge_colors) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
emm_noGM

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\emm_noGM.png", 
       plot = emm_noGM, 
       width = 8, 
       height = 5, 
       dpi = 300,
       limitsize = FALSE)
```

## Temporal grain comparison: Month vs. Phen_Period vs. Cal_Season

The models above use `Month` (calendar month) as the temporal grouping. `Phen_Period` (existing, `Study.Phase`-based, 4 levels) and `Cal_Season` (new, Lacy et al. 2019, 2 levels) are both alternative temporal groupings of the same data -- run here as separate, explicitly-labeled models, not as replacements for the Month-based model above. Comparing all three tells you whether the marsh_edge_type x time pattern is robust to how "time" gets grouped, or whether it's sensitive to that choice -- itself a useful thing to know before deciding which grain to lead with in the defense.

```{r, edge phen period model}
# SENSITIVITY check (Rule 2) -- edge_phenperiod_model_glmm below is
# primary. Same (1|Site) structure as edge_time_model above, swapping
# Month for Phen_Period.

edge_phenperiod_model <- lmer(
  Flux ~ marsh_edge_type * Phen_Period + (1|Site),
  data = master_tier5_clean
)
summary(edge_phenperiod_model)
anova(edge_phenperiod_model)
```

```{r, edge phen period glmm}
# PRIMARY model for the edge-type x Phen_Period comparison (Rule 2).

edge_phenperiod_model_glmm <- glmmTMB(
  Flux ~ marsh_edge_type * Phen_Period + (1|Site),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)
summary(edge_phenperiod_model_glmm)
Anova(edge_phenperiod_model_glmm, type = "II")

sim_res_edge_phenperiod <- simulateResiduals(edge_phenperiod_model_glmm)
plot(sim_res_edge_phenperiod)
testDispersion(sim_res_edge_phenperiod)

AIC(edge_time_model_glmm, edge_phenperiod_model_glmm)
```

```{r, edge phen period emm plot}
emm_edge_phenperiod <- emmeans(edge_phenperiod_model_glmm, ~ marsh_edge_type * Phen_Period, type = "response")
emm_edge_phenperiod_df <- as.data.frame(emm_edge_phenperiod) %>%
  rename(emmean = response, lower.CL = asymp.LCL, upper.CL = asymp.UCL)

emm_phenperiod = ggplot(emm_edge_phenperiod_df, aes(x = Phen_Period, y = emmean, color = marsh_edge_type, group = marsh_edge_type)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(0.3)) +
  labs(
    title = str_wrap("Estimated Marginal Means of Flux by Marsh Edge Type Across Phenological Period", width = 40),
    y = "Estimated Flux", x = "Phenological Period", color = "Edge Type"
  ) +
  theme_bw() +
  scale_color_manual(values = marsh_edge_colors) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
emm_phenperiod

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\emm_phenperiod.png",
       plot = emm_phenperiod,
       width = 8, height = 5, dpi = 300, limitsize = FALSE)
```

```{r, edge season model}
# SENSITIVITY check (Rule 2) -- edge_season_model_glmm below is
# primary. Same (1|Site) structure as edge_time_model above, swapping
# Month for Cal_Season (Lacy et al. 2019: Nov-Feb Winter, Mar-Oct Summer).

edge_season_model <- lmer(
  Flux ~ marsh_edge_type * Cal_Season + (1|Site),
  data = master_tier5_clean
)
summary(edge_season_model)
anova(edge_season_model)
```

```{r, edge season glmm}
# PRIMARY model for the edge-type x Cal_Season comparison (Rule 2).

edge_season_model_glmm <- glmmTMB(
  Flux ~ marsh_edge_type * Cal_Season + (1|Site),
  family = Gamma(link = "log"),
  data = master_tier5_clean
)
summary(edge_season_model_glmm)
Anova(edge_season_model_glmm, type = "II")

sim_res_edge_season <- simulateResiduals(edge_season_model_glmm)
plot(sim_res_edge_season)
testDispersion(sim_res_edge_season)

AIC(edge_time_model_glmm, edge_season_model_glmm)
```

```{r, edge season emm plot}
emm_edge_season <- emmeans(edge_season_model_glmm, ~ marsh_edge_type * Cal_Season, type = "response")
emm_edge_season_df <- as.data.frame(emm_edge_season) %>%
  rename(emmean = response, lower.CL = asymp.LCL, upper.CL = asymp.UCL)

emm_season = ggplot(emm_edge_season_df, aes(x = Cal_Season, y = emmean, color = marsh_edge_type, group = marsh_edge_type)) +
  geom_point(size = 3, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, position = position_dodge(0.3)) +
  labs(
    title = str_wrap("Estimated Marginal Means of Flux by Marsh Edge Type Across Cal_Season (Lacy et al. 2019)", width = 40),
    y = "Estimated Flux", x = "Cal_Season", color = "Edge Type"
  ) +
  theme_bw() +
  scale_color_manual(values = marsh_edge_colors) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
emm_season

ggsave("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\images\\rq_1\\emm_season.png",
       plot = emm_season,
       width = 8, height = 5, dpi = 300, limitsize = FALSE)
```

**On the AIC comparisons above:** `edge_time_model_glmm`, `edge_phenperiod_model_glmm`, and `edge_season_model_glmm` are all fit on the same `master_tier5_clean` (same N), so unlike the averaged-vs-replicate comparison earlier in this project, `AIC()` across these three *is* a valid comparison -- it tells you which temporal grain the data actually supports best, not an artifact of differing sample size. A lower AIC for `Cal_Season` over `Month`, for instance, would be a legitimate argument for leading with the coarser, literature-backed grouping in the defense rather than the finer one.

```{r}

# Replace `Interval_Length` with whatever your actual column is called
# (Date.removed.from.field - Date.placed.in.field, in days) in the code
# that built rq_plots.

interval_summary <- master_tier5 %>%
  mutate(Interval_Length = as.numeric(Date.removed.from.field - Date.placed.in.field))

# Statistical threshold rather than an eyeballed cutoff -- flag anything
# above Q3 + 1.5*IQR (standard outlier rule), OR anything you and Matt
# agree is operationally "too long to trust" (e.g., >35 days). Pick one;
# shown here with both so you can compare.
q <- quantile(interval_summary$Interval_Length, c(0.25, 0.75), na.rm = TRUE)
iqr_cutoff <- q[2] + 1.5 * (q[2] - q[1])

long_intervals <- interval_summary %>%
  filter(Interval_Length > iqr_cutoff | Interval_Length > 35) %>%
  select(Site, Month, Study.Phase, Date.placed.in.field, Date.removed.from.field, Interval_Length) %>%
  arrange(Date.removed.from.field)

long_intervals
```

```{r}

# ---- 2. Exclusion list -----------------------------------------------------

# Option A (commented out for now) -- site-specific exclusion, pulled
# directly from long_intervals. Revisit this once the broad test below
# is done and you want a more targeted version.
# exclude_site_phase <- long_intervals %>%
#   distinct(Site, Study.Phase)

# Option B (active) -- drop December everywhere, for a first-pass sensitivity check
december_phases <- master_tier5 %>%
  filter(Month == "Dec") %>%
  distinct(Study.Phase) %>%
  pull(Study.Phase)

december_phases

# ---- 3. Build the sensitivity tier ----------------------------------------
master_tier5_sensitivity <- filter_flux(elev_z_wind_slope_edge_flux, "all", "full") %>%
  filter(!(Study.Phase %in% december_phases)) %>%
  # anti_join(exclude_site_phase, by = c("Site", "Study.Phase")) %>%   # swap in for the site-specific version later
  average_replicates() %>%
  filter(n_reps > 0)

# ---- 4. Run both scenarios -------------------------------------------------
general_with    <- make_flux_table(master_tier5,             title = "Including December")
general_without <- make_flux_table(master_tier5_sensitivity, title = "Excluding December")

phase_with    <- make_flux_table(master_tier5,             group_vars = "Study.Phase",
                                 title = "By Study Phase — including December",
                                 group_labels = list(Study.Phase = "Study Phase"))
phase_without <- make_flux_table(master_tier5_sensitivity, group_vars = "Study.Phase",
                                 title = "By Study Phase — excluding December",
                                 group_labels = list(Study.Phase = "Study Phase"))

# ---- 5. Comparison tables --------------------------------------------------
overall_comparison <- bind_rows(
  general_with$data    %>% mutate(Scenario = "Including December", .before = 1),
  general_without$data %>% mutate(Scenario = "Excluding December", .before = 1)
)

phase_comparison <- bind_rows(
  phase_with$data    %>% mutate(Scenario = "Including December", .before = 1),
  phase_without$data %>% mutate(Scenario = "Excluding December", .before = 1)
) %>%
  arrange(Study.Phase, Scenario)

overall_comparison
phase_comparison
```

```{r}

# ---- Exclusion list: every Site + Study.Phase combo flagged as a long interval
exclude_site_phase <- long_intervals %>%
  distinct(Site, Study.Phase)

exclude_site_phase   # check this before it drives anything below

# ---- Build the sensitivity tier, excluding all flagged combos -----------
master_tier5_sensitivity_all <- filter_flux(elev_z_wind_slope_edge_flux, "all", "full") %>%
  anti_join(exclude_site_phase, by = c("Site", "Study.Phase")) %>%
  average_replicates() %>%
  filter(n_reps > 0)

nrow(master_tier5)
nrow(master_tier5_sensitivity_all)   # should be smaller than master_tier5, and smaller
# than master_tier5_sensitivity (December-only)


```

```{r}

phase_flux_by_scenario <- bind_rows(
  master_tier5                  %>% mutate(Scenario = "Including December"),
  master_tier5_sensitivity      %>% mutate(Scenario = "Excluding December only"),
  master_tier5_sensitivity_all  %>% mutate(Scenario = "Excluding all flagged intervals")
) %>%
  mutate(
    Scenario = factor(Scenario, levels = c("Including December",
                                            "Excluding December only",
                                            "Excluding all flagged intervals")), 
                       Study.Phase = factor(Study.Phase, levels = paste0("M", 1:12))
)
    # Force correct M1-M12 order regardless of what came in from bind_rows


phase_flux_boxplot_dodged <- ggplot(phase_flux_by_scenario,
                                     aes(x = Study.Phase, y = Flux, fill = Scenario)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  scale_y_log10() +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Study Phase (Month)", y = expression(Log[10]~Flux~(g/m^2/day)),
       title = "Flux by Study Phase Across Data-Exclusion Scenarios",
       fill = "Scenario")

phase_flux_boxplot_dodged

ggsave(file.path(knitr::opts_knit$get("root.dir"), "phase_flux_by_scenario_dodged.png"),
       plot = phase_flux_boxplot_dodged,
       width = 13,
       height = 7,
       dpi = 300,
       limitsize = FALSE)



site_point_plot_grid <- ggplot(phase_flux_by_scenario,
                                aes(x = Study.Phase, y = Flux, color = Site)) +
  geom_point(alpha = 0.5, size = 1.5, position = position_jitter(width = 0.2)) +
  scale_y_log10(labels = label_number(accuracy = 1)) +
  scale_color_brewer(palette = "Dark2") +
  facet_grid(Scenario ~ Site) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
        strip.text.y = element_text(face = "bold", size = 10, angle = 0),
        strip.text.x = element_text(face = "bold", size = 11),
        panel.grid.minor = element_blank()) +
  labs(x = "Study Phase (Month)", y = expression(Log[10]~Flux~(g/m^2/day)),
       title = "Flux by Study Phase and Site\nAcross Data-Exclusion Scenarios")

site_point_plot_grid


ggsave(file.path(knitr::opts_knit$get("root.dir"), "site_point_grid.png"),
       plot = site_point_plot_grid,
       width = 15,
      height = 11,
       dpi = 300,
       limitsize = FALSE)
```

```{r}

general_all <- make_flux_table(master_tier5_sensitivity_all, title = "Excluding all flagged long intervals")

phase_all <- make_flux_table(master_tier5_sensitivity_all, group_vars = "Study.Phase",
                             title = "By Study Phase — excluding all flagged long intervals",
                             group_labels = list(Study.Phase = "Study Phase"))

overall_comparison_3way <- bind_rows(
  general_with$data       %>% mutate(Scenario = "Including all",                .before = 1),
  general_without$data    %>% mutate(Scenario = "Excluding December only",      .before = 1),
  general_all$data        %>% mutate(Scenario = "Excluding all flagged intervals", .before = 1)
)

flagged_phases <- unique(long_intervals$Study.Phase)

phase_comparison_3way <- bind_rows(
  phase_with$data %>% mutate(Scenario = "Including all", .before =1), 
  phase_without$data %>% 
    filter(Study.Phase %in% december_phases) %>% 
    mutate(Scenario = "Excluding December only", .before =1),  
  phase_all$data %>% 
    filter(Study.Phase %in% flagged_phases) %>%
    mutate(Scenario = "Excluding all flagged intervals", .before = 1)) %>%
   mutate(Study.Phase = factor(Study.Phase, levels = paste0("M",1:12), ordered = TRUE)) %>%
  arrange(Study.Phase, Scenario)

overall_comparison_3way
phase_comparison_3way
```

```{r, december deployment dates table}


library(gt)

december_intervals <- master_tier5 %>%
  mutate(Interval_Length = as.numeric(Date.removed.from.field - Date.placed.in.field)) %>%
  filter(Month == "Dec") %>%
  select(Site, Study.Phase, Date.placed.in.field, Date.removed.from.field, Interval_Length) %>%
  arrange(Site, Date.placed.in.field)

december_table <- december_intervals %>%
  gt() %>%
  tab_header(title = "December Deployment and Removal Dates") %>%
  cols_label(
    Site = "Site",
    Study.Phase = "Study Phase",
    Date.placed.in.field = "Date Placed",
    Date.removed.from.field = "Date Removed",
    Interval_Length = "Interval (days)"
  ) %>%
  fmt_number(columns = Interval_Length, decimals = 1)

december_table

gtsave(
  december_table,
  filename = file.path(knitr::opts_knit$get("root.dir"), "december_deployment_dates.png")
)

```

```{r}
overall_comparison_3way_table <- overall_comparison_3way %>%
  gt() %>%
  tab_header(title = "Overall Flux Comparison Across December-Exclusion Scenarios") %>%
  fmt_number(columns = where(is.numeric), decimals = 2)

overall_comparison_3way_table

phase_comparison_3way_table <-phase_comparison_3way %>% 
  gt() %>% 
  tab_header(title = "Flux by Study Phase Across Study Phases with Long Intervals > 35 Days") %>%
  fmt_number(columns = where(is.numeric), decimals = 2)

phase_comparison_3way_table

gtsave(
  overall_comparison_3way_table,
  filename = file.path(knitr::opts_knit$get("root.dir"), "overall_comparison_3way.png")
)

gtsave(
 phase_comparison_3way_table,
  filename = file.path(knitr::opts_knit$get("root.dir"), "phase_comparison_3way.png")
)

```
