---
title: "Sediment Deposition Rates Across Distances and Plant Dimensions\n In Two Marshes of San Pablo Bay, CA"
author: "Savannah K. Miller"
date: "2025-03-21"
output: pdf_document
---

#### Dataset Card Source: (<https://github.com/Smillersfsu/DepositionSFB/blob/0a6b8b2033b76344b3b1574f88ab81aa244e5210/DatasetCard.md>)

## Background

Sediment is the structural foundation of the salt marsh ecosystem where endemic species live and acquire resources (Levin et al. 2001; Janousek et al. 2017). The salt marsh ecosystem is host to many endemic species that are specially adapted to the conditions of the salt marsh (Dunson and Travis 1994; Schile et al. 2011; Moffett, Robinson, and Gorelick 2010; Rankin et al. 2023), and the very sediment they walk and live within is in short supply (Schoellhamer and Marineau, n.d.). Salt marshes play strong roles in dampening impacts of atmospheric rivers and extreme weather events, offering a protection to coastal communities(Smolders et al. 2015; Castagno et al. 2022; Lee and Nepf 2024; Taylor-Burns et al. 2024). Already, there has been a long human history of disturbance to existing salt marshes, which puts ecological and human communities at severe risk(Gedan, Silliman, and Bertness 2009). Under these circumstances, the few salt marshes left are at risk of elimination due to drowning if action is not taken and the benefits of this ecosystem could be forever lost.(Endris et al. 2024).

Across the state, salt marshes are decreasing in number and area due to a variety of factors (Endris et al. 2024). This unfortunate circumstance paired with changing water levels along the coastline increases the sense of urgency to understand, conserve, and restore California’s salt marshes. To assist in raising the elevation of the salt marshes to adapt to natural hazards, we need to understand the fundamental processes of how sediment accumulates within the salt marsh over a brief period (short-term deposition) (Fagherazzi 2013; Houttuijn Bloemendaal et al. 2021; Vandenbruwaene et al. 2011). In this study, I will measure short-term deposition rates across marsh edges of multiple marshes within a single estuary in San Pablo Bay of the San Francisco Bay Estuary and comparing those rates to measurements of the surrounding marsh landscape. This study will provide information to ongoing restoration efforts of salt marshes in San Pablo Bay.(Callaway 2024; University of San Francisco et al. 2011; K. M. Thorne et al. 2019; Haltiner et al. 1996) My study provides novel insights for land managers to characterize their marshes edges to adapt to natural hazards and enhance community resiliency (Reed et al. 1999).

In this study, I will measure short-term deposition rates on multiple marshes within the San Francisco Bay Estuary and will compare those rates to measurements of the surrounding marsh landscape. This study builds on the work of Karen Thorne and U.S. Geological Survey (2025; Inter-and Intra-Annual Sediment Dynamics in Two Tidal Marshes: Deposition, Accretion, and Vegetation Data - ScienceBase-Catalog) to understand differences in sediment deposition on marshes with different edge characteristics, vegetation densities and geographic settings. Moreover, I will measure variables such as elevation, vegetation structure, and observations of marsh edge traits and comparing to rates of short-term deposition measured to understand patterns between variables. Other environmental variables such as season at time of collection and tide level at time of collection will be recorded and considered in analyses.

## Research Objective

This study will measure short-term deposition rates of sediment across the exposed (bay-adjacent) salt marshes in the northern part of the San Francisco Bay Estuary of the San Pablo Bay. I will compare these rates to biological structures of each salt marsh.

### Research Aim:

Marsh characteristics of the vegetation also are one of the first interactions with the waves bay ward and along the marsh edge. These vegetation characteristics will be measured at each marsh and compared with the rates of sediment deposition measured over the course of a year’s spring-neap tidal cycles.

### Expected Outcome:

Deposition rates vary significantly with characteristics of the marsh vegetation, specifically maximum height.

### Null Hypothesis:

Deposition rates of sediment do not vary with differences in height of Salicornia pacificia (Pickleweed).

### Alternate hypothesis

Deposition rates of sediment do vary with differences in height of Salicornia pacifica (Pickleweed).

## Data analysis Plan

First, I played with both the vegetation and sediment data sets to understand patterns that each of those explanatory variables had with the landscape and abiotic factors, such as distance from the water source ("Distance"), weather factors ("Season"), time factors ("Year" and "Month"), and location ("Site").

As the sample distribution for the sediment flux measurements was not normally distributed, which was my main independent variable, I did not run a two-sampled t-test.

I ran a general linear mixed model with both sites data and sediment deposition versus the maximum plant height. I followed this by running a general linear mixed model for each site with those same explanatory variables to compare the AIC values and find the best fit, with distance from the marsh edge as an additional predictor. Within these models, I accounted for random effects such as "Year", "Season" and "Site".

```{r, Load necessary packages}

# data import and database management 
# reading files, reshaping data, and efficient data handling 

#######################################################

library(readxl) # import excel spreadsheets
library(data.table) #quick manipulation of large datasets
library(stringr) # text and string processing 
library(lubridate) # data and time handling 
library(writexl) # exports editted databases into excel files

#######################################################

# General data science workflow 
# Collection of packages for data manipulation and plotting

#######################################################

library(tidyverse)

#######################################################

#Statistical analyses and mixed models 
# Hypothesis testing, summaries, and model fitting 

#######################################################

library(gt) # displaying formatted tables for statistics 
library(rstatix) # statistical tests and assumption checking
library(lmerTest) # p-values and tests for lme4 models 
library(lme4) # linear and generalized mixed-effects models
library(ggeffects) # predicted values and marginal effects 

#######################################################

# Data visualization 
# Plot creation, figure assembly, and publication formatting 

#######################################################

library(patchwork) #comvine ggplots into multi-panel figures
library(cowplot) # figure arrangement and plot annotations
library(rcartocolor) # color palettes
library(ggh4x) # advanced ggplot extensions
library(scales) # axis labels, formatting, and scaling 

#######################################################
# Spatial data and mapping 
# GIS data handling and map visualization
#######################################################

library(sf) #spatial vector data
library(ggmap) #basemaps and geographic visualization
library(ggspatial) #spatial annotations and map elements
library(leaflet) #interactive web maps

#######################################################

#Tables and report outputs 
# Interactive tables and report presentation 

#######################################################

library(reactable) #interactive tables for reports

#######################################################
# R markdown and document rendering 
# PDF generation and LaTeX support 
#######################################################

library(tinytex) #LaTeX engine for PDF knitting


```

```{r, set working directory}

knitr::opts_knit$set(root.dir = "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data")

```

```{r, cleaning data}
# load data
thesis_veg <- read.csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\thesis_veg.csv") 

usgs_veg <- read.csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_SFB_1m_vegetation_data.csv") 


thesis_veg <- thesis_veg %>% mutate(AvgHt = ifelse(is.na(AvgHt), 0, AvgHt))
thesis_veg <- thesis_veg %>% mutate(MaxHt = ifelse(is.na(MaxHt), 0, MaxHt))
thesis_veg$Season <- str_replace_all(thesis_veg$Season, " ", "")
thesis_veg$Plot_ID <- str_replace_all(thesis_veg$Plot_ID, " ", "")
thesis_veg$Year <- str_replace_all(thesis_veg$Year, " ", "")
thesis_veg$Month <- str_replace_all(thesis_veg$Month, " ", "")
thesis_veg$Year <- as.character(thesis_veg$Year)
thesis_veg$Cover <-as.numeric(thesis_veg$Cover)
thesis_veg <- thesis_veg %>%
  separate(Plot_ID, into = c("Tran", "Dist"), sep = "-", remove = FALSE )
thesis_veg$Dist <- as.numeric(thesis_veg$Dist)

usgs_veg <- usgs_veg %>% mutate(AvgHt = ifelse(is.na(AvgHt), 0, AvgHt))
usgs_veg <- usgs_veg %>% mutate(MaxHt = ifelse(is.na(MaxHt), 0, MaxHt))
usgs_veg$Season <- str_replace_all(usgs_veg$Season, " ", "")
usgs_veg$Plot.ID <- str_replace_all(usgs_veg$Plot.ID, " ", "")
usgs_veg$Year <- str_replace_all(usgs_veg$Year, " ", "")
usgs_veg$Month <- str_replace_all(usgs_veg$Month, " ", "")
usgs_veg$Year <- as.character(usgs_veg$Year)
usgs_veg$Cover <-as.numeric(usgs_veg$Cover)
usgs_veg <-usgs_veg %>%
  mutate(
    Tran = str_extract(Plot.ID, "^[A-Za-z]+"), 
    Dist = as.numeric(str_extract(Plot.ID, "[0-9.]+$"))
  )

str(thesis_veg)
str(usgs_veg)


```



```{r, veg and SAPA combined dataframe}

sed_SAPA <- sed %>% left_join(veg, by = c("Site", "Season", "Distance", "Transect", "Year", "Month"))
sed_SAPA <- sed_SAPA %>%
  mutate(FluxAvgRep = ifelse(is.na(FluxAvgRep), 0, FluxAvgRep),
        Species = ifelse(is.na(Species), 0, Species), 
        AvgHt = ifelse(is.na(AvgHt), 0, AvgHt),
        MaxHt = ifelse(is.na(MaxHt), 0, MaxHt), 
        Cover = ifelse(is.na(Cover), 0, Cover))
sed_SAPA <- sed_SAPA %>%
  select(-Date)

```



```{r, focus on SAPA}

SAPA_usgs <- usgs_veg %>% 
  filter(Species == "SAPA")

SAPA_thesis <- thesis_veg %>% 
  filter(Species ==  "SAPA")

SAPA_dist_mx_usgs <- ggplot(data=SAPA_usgs, aes(x=Dist, y=MaxHt, color=Site)) + geom_point() +  facet_wrap(~ Season)
SAPA_dist_mx_usgs

SAPA_dist_avg_usgs <- ggplot(data=SAPA_usgs, aes(x=Dist, y=AvgHt, color=Site)) + geom_point() + facet_wrap(~ Season)
SAPA_dist_avg_usgs

SAPA_dist_c_usgs <- ggplot(data=SAPA_usgs, aes(x=Dist, y=Cover, color=Site)) + geom_point() +facet_wrap(~ Season)
SAPA_dist_c_usgs

#thesis 

SAPA_dist_mx_thesis <- ggplot(data=SAPA_thesis, aes(x=Dist, y=MaxHt, color=Site)) + geom_point() +  facet_wrap(~ Season)
SAPA_dist_mx_thesis

SAPA_dist_avg_thesis <- ggplot(data=SAPA_thesis, aes(x=Dist, y=AvgHt, color=Site)) + geom_point() + facet_wrap(~ Season)
SAPA_dist_avg_thesis

SAPA_dist_c_thesis <- ggplot(data=SAPA_thesis, aes(x=Dist, y=Cover, color=Site)) + geom_point() +facet_wrap(~ Season)
SAPA_dist_c_thesis
```



```{r, focus on SPFO}

SPFO_usgs <- usgs_veg %>% 
  filter(Species == "SPFO")

SPFO_thesis <- thesis_veg %>% 
  filter(Species ==  "SPFO")

SPFO_dist_mx_usgs <- ggplot(data=SPFO_usgs, aes(x=Dist, y=MaxHt, color=Site)) + geom_point() +  facet_wrap(~ Season)
SPFO_dist_mx_usgs

SPFO_dist_avg_usgs <- ggplot(data=SPFO_usgs, aes(x=Dist, y=AvgHt, color=Site)) + geom_point() + facet_wrap(~ Season)
SPFO_dist_avg_usgs

SPFO_dist_c_usgs <- ggplot(data=SPFO_usgs, aes(x=Dist, y=Cover, color=Site)) + geom_point() +facet_wrap(~ Season)
SPFO_dist_c_usgs

#thesis 

SPFO_dist_mx_thesis <- ggplot(data=SPFO_thesis, aes(x=Dist, y=MaxHt, color=Site)) + geom_point() +  facet_wrap(~ Season)
SPFO_dist_mx_thesis

SPFO_dist_avg_thesis <- ggplot(data=SPFO_thesis, aes(x=Dist, y=AvgHt, color=Site)) + geom_point() + facet_wrap(~ Season)
SPFO_dist_avg_thesis

SPFO_dist_c_thesis <- ggplot(data=SPFO_thesis, aes(x=Dist, y=Cover, color=Site)) + geom_point() +facet_wrap(~ Season)
SPFO_dist_c_thesis
```


```{r, visualizing relationships of the sediment data to explanatory variables}

sed_space <-ggplot(data=sed_SAPA, aes(x=Distance, y=FluxAvgRep, color=Site)) +geom_point() + geom_jitter(width = 0.2, height = 0.2, alpha = 0.6) + theme_minimal() +
 coord_cartesian(ylim = c(0, 357)) + facet_wrap(~ Season)
sed_space

sed_time <-ggplot(data=sed_SAPA, aes(x=Season, y=FluxAvgRep, color=Site)) + geom_point() + geom_jitter(width = 0.2, height = 0.2, alpha = 0.6) + theme_minimal() +
 coord_cartesian(ylim = c(0, 357))
sed_time

sed_year <- ggplot(sed_SAPA, aes(x= Season, y= FluxAvgRep, color= Site)) +geom_point() + facet_wrap(~Year)
sed_year

sed_season_site = ggplot(sed, aes(x = Season, y = FluxAvgRep, fill = Site)) +
  facet_wrap(~Distance) +
  geom_col(position = "dodge") +
  scale_y_log10() +
  labs(x = "Season", y = "Log Flux of Sediment Deposition") 
sed_season_site

sed_Siteseason <- ggplot(sed_SAPA, aes(x= Site, y=FluxAvgRep, color = Season)) +geom_point() + geom_boxplot()
sed_Siteseason

sed_season = ggplot(sed, aes(x = Site, y = FluxAvgRep, fill = Season)) + 
  geom_col(position = "dodge") +
  scale_y_log10() +
  labs(x = "Site", y = "Log Flux of Sediment Deposition") 
sed_season

sed_Dist = ggplot(sed, aes(x = Distance, y = FluxAvgRep, color = Site)) + 
  geom_point(aes(x = Distance, y = FluxAvgRep), data = sed, position = position_jitter(w = 0.05, h = 0)) +
  #scale_y_log10() + 
   geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance", y = "Log Flux of Sediment Deposition", 
       title = "Sediment Deposition Rates Across Distances")
sed_Dist
```



```{r}
# usgs 

SAPA_Season_mxht_usgs <- ggplot(SAPA_usgs, aes(x=MaxHt,fill=Season)) + geom_boxplot()
SAPA_Season_mxht_usgs

SAPA_Season_avght_usgs <-ggplot(SAPA_usgs, aes(x=AvgHt,fill=Season)) + geom_boxplot()
SAPA_Season_avght_usgs

SAPA_Season_cover_usgs <- ggplot(SAPA_usgs, aes(x=Cover, fill = Season)) + geom_boxplot()
SAPA_Season_cover_usgs

SAPA_Site_cover_usgs <- ggplot(SAPA_usgs, aes(x= Cover, fill = Site)) + geom_boxplot()
SAPA_Site_cover_usgs

SAPA_Site_mxht_usgs <- ggplot(SAPA_usgs, aes(x = MaxHt, fill = Site)) + geom_boxplot()
SAPA_Site_mxht_usgs

SAPA_Site_avght_usgs <- ggplot(SAPA_usgs, aes(x = AvgHt, fill = Site)) + geom_boxplot()
SAPA_Site_avght_usgs

#check the data for normality, MaxHt 
SAPA_Maxht_Normality_usgs <-ggplot(SAPA_usgs, aes(x=MaxHt)) + geom_histogram(binwidth = 0.5)
SAPA_Maxht_Normality_usgs

#check the data for normality, AvgHt  
SAPA_Avght_Normality_usgs <-ggplot(SAPA_usgs, aes(x=AvgHt)) + geom_histogram(binwidth = 0.5)
SAPA_Avght_Normality_usgs

#check the data for normality, Cover 
SAPA_Cover_Normality_usgs <-ggplot(SAPA_usgs, aes(x=Cover)) + geom_histogram(binwidth = 0.5)
SAPA_Cover_Normality_usgs
```










```{r}
# thesis

SAPA_Season_mxht_thesis <- ggplot(SAPA_thesis, aes(x=MaxHt,fill=Season)) + geom_boxplot()
SAPA_Season_mxht_thesis

SAPA_Season_avght_thesis <-ggplot(SAPA_thesis, aes(x=AvgHt,fill=Season)) + geom_boxplot()
SAPA_Season_avght_thesis

SAPA_Season_cover_thesis <- ggplot(SAPA_thesis, aes(x=Cover, fill = Season)) + geom_boxplot()
SAPA_Season_cover_thesis

SAPA_Site_cover_thesis <- ggplot(SAPA_thesis, aes(x= Cover, fill = Site)) + geom_boxplot()
SAPA_Site_cover_thesis

SAPA_Site_mxht_thesis <- ggplot(SAPA_thesis, aes(x = MaxHt, fill = Site)) + geom_boxplot()
SAPA_Site_mxht_thesis

SAPA_Site_avght_thesis <- ggplot(SAPA_thesis, aes(x = AvgHt, fill = Site)) + geom_boxplot()
SAPA_Site_avght_thesis

#check the data for normality, MaxHt 
SAPA_Maxht_Normality_thesis <-ggplot(SAPA_thesis, aes(x=MaxHt)) + geom_histogram(binwidth = 0.5)
SAPA_Maxht_Normality_thesis

#check the data for normality, AvgHt  
SAPA_Avght_Normality_thesis <-ggplot(SAPA_thesis, aes(x=AvgHt)) + geom_histogram(binwidth = 0.5)
SAPA_Avght_Normality_thesis

#check the data for normality, Cover 
SAPA_Cover_Normality_thesis <-ggplot(SAPA_thesis, aes(x=Cover)) + geom_histogram(binwidth = 0.5)
SAPA_Cover_Normality_thesis
```











```{r, SPFO-checking the range of data for the vegetation dataset}

# usgs 

SPFO_Season_mxht_usgs <- ggplot(SPFO_usgs, aes(x=MaxHt,fill=Season)) + geom_boxplot()
SPFO_Season_mxht_usgs

SPFO_Season_avght_usgs <-ggplot(SPFO_usgs, aes(x=AvgHt,fill=Season)) + geom_boxplot()
SPFO_Season_avght_usgs

SPFO_Season_cover_usgs <- ggplot(SPFO_usgs, aes(x=Cover, fill = Season)) + geom_boxplot()
SPFO_Season_cover_usgs

SPFO_Site_cover_usgs <- ggplot(SPFO_usgs, aes(x= Cover, fill = Site)) + geom_boxplot()
SPFO_Site_cover_usgs

SPFO_Site_mxht_usgs <- ggplot(SPFO_usgs, aes(x = MaxHt, fill = Site)) + geom_boxplot()
SPFO_Site_mxht_usgs

SPFO_Site_avght_usgs <- ggplot(SPFO_usgs, aes(x = AvgHt, fill = Site)) + geom_boxplot()
SPFO_Site_avght_usgs

#check the data for normality, MaxHt 
SPFO_Maxht_Normality_usgs <-ggplot(SPFO_usgs, aes(x=MaxHt)) + geom_histogram(binwidth = 0.5)
SPFO_Maxht_Normality_usgs

#check the data for normality, AvgHt  
SPFO_Avght_Normality_usgs <-ggplot(SPFO_usgs, aes(x=AvgHt)) + geom_histogram(binwidth = 0.5)
SPFO_Avght_Normality_usgs

#check the data for normality, Cover 
SPFO_Cover_Normality_usgs <-ggplot(SPFO_usgs, aes(x=Cover)) + geom_histogram(binwidth = 0.5)
SPFO_Cover_Normality_usgs

# thesis

SPFO_Season_mxht_thesis <- ggplot(SPFO_thesis, aes(x=MaxHt,fill=Season)) + geom_boxplot()
SPFO_Season_mxht_thesis

SPFO_Season_avght_thesis <-ggplot(SPFO_thesis, aes(x=AvgHt,fill=Season)) + geom_boxplot()
SPFO_Season_avght_thesis

SPFO_Season_cover_thesis <- ggplot(SPFO_thesis, aes(x=Cover, fill = Season)) + geom_boxplot()
SPFO_Season_cover_thesis

SPFO_Site_cover_thesis <- ggplot(SPFO_thesis, aes(x= Cover, fill = Site)) + geom_boxplot()
SPFO_Site_cover_thesis

SPFO_Site_mxht_thesis <- ggplot(SPFO_thesis, aes(x = MaxHt, fill = Site)) + geom_boxplot()
SPFO_Site_mxht_thesis

SPFO_Site_avght_thesis <- ggplot(SPFO_thesis, aes(x = AvgHt, fill = Site)) + geom_boxplot()
SPFO_Site_avght_thesis

#check the data for normality, MaxHt 
SPFO_Maxht_Normality_thesis <-ggplot(SPFO_thesis, aes(x=MaxHt)) + geom_histogram(binwidth = 0.5)
SPFO_Maxht_Normality_thesis

#check the data for normality, AvgHt  
SPFO_Avght_Normality_thesis <-ggplot(SPFO_thesis, aes(x=AvgHt)) + geom_histogram(binwidth = 0.5)
SPFO_Avght_Normality_thesis

#check the data for normality, Cover 
SPFO_Cover_Normality_thesis <-ggplot(SPFO_thesis, aes(x=Cover)) + geom_histogram(binwidth = 0.5)
SPFO_Cover_Normality_thesis
```





```{r, Fit GLM with Poisson family}

sedvegglm <-glm(FluxAvgRep~MaxHt, family = "poisson", data=sed_SAPA)
sedvegglm

sedvegglm_AIC_text <- paste("AIC: 7914")
sedvegglm_AIC_text

sedvegglm_coefficient_MaxHt <- paste("Coefficient for Max Height: -0.01325 ")
sedvegglm_coefficient_MaxHt

```

```{r, Plot model diagnostics}

plot(sedvegglm)
```

```{r, glmm trial and error with random effects}

library(glmmTMB)
sed_glmm = glmmTMB(FluxAvgRep~MaxHt + Distance + (1|Site) + (1|Season) + (1|Year), family = poisson(), data=sed_SAPA)
summary(sed_glmm)

sed_glmm2 = glmmTMB(FluxAvgRep~MaxHt + Distance + (1|Site) + (1|Season), family = poisson(), data=sed_SAPA)
summary(sed_glmm2)

sed_glmm3 = glmmTMB(FluxAvgRep~MaxHt + Distance + (1|Site), family = poisson(), data=sed_SAPA)
summary(sed_glmm3)

sed_glmm4 = glmmTMB(FluxAvgRep~MaxHt + Distance + (1|Site) + (1|Year), family = poisson(), data=sed_SAPA)
summary(sed_glmm4)

sed_glmm5 = glmmTMB(FluxAvgRep~MaxHt + Distance + (1|Year) + (1|Season), family = poisson(), data=sed_SAPA)
summary(sed_glmm5)

AIC(sed_glmm)
AIC(sed_glmm, sed_glmm2, sed_glmm3, sed_glmm4, sed_glmm5)

```

```{r, graphing the relationships}

sed_MxHt_season <- ggplot(sed_SAPA, aes(x = MaxHt, y = FluxAvgRep, color = Season)) +
  geom_point(position = position_jitter(w = 0.05, h = 0)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10() +
  labs(
    x = expression("Maximum Height of Pickleweed " * italic("(Salicornia pacifica)")),
    y = "Log Flux of Sediment Deposition",
    title = "Sediment Deposition Rates Across Differences in Maximum Plant Height\n Between Two Seasons"
  )

sed_MxHt_season

sed_DistPlant <- ggplot(veg, aes(x = Distance, y = MaxHt, color = Site)) + 
  geom_point(position = position_jitter(w = 0.05, h = 0)) +
  scale_y_log10() + 
  labs(
    x = "Distance from Marsh Edge",
    y = expression("Maximum Height of Pickleweed " * italic("(Salicornia pacifica)")),
    title = "Sediment Deposition Rates Across Plant Structure\n(Maximum Height)"
  )

sed_DistPlant

```

## Summary of Results/Conclusions

```{r, final figure}

sed_MxHt_site <- ggplot(sed_SAPA, aes(x = MaxHt, y = FluxAvgRep, color = Site)) +
  geom_point(position = position_jitter(w = 0.05, h = 0)) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10() +
  labs(
    x = expression("Maximum Height of Pickleweed " * italic("(Salicornia pacifica)")),
    y = "Log Flux of Sediment Deposition",
    title = "Sediment Deposition Rates Across Differences in Maximum Plant Height\n Between Two Marshes"
  )
sed_MxHt_site

```

Within this graph, sediment deposition is marked with a log function as there were some outliers within the data set, specifically at the San Pablo site that made it difficult to observe patterns in the data.

The graph above and the GLMM outputs demonstrates that maximum height of Pickleweed does play a role in sediment deposition, though site specific differences play a role in influencing those relationships. Season and Year also did play a role in the strength of these predictor variables, however since some of the interest of this study is based off of replication across site, this was the final graph that I chose.

Based off of the conclusions of the linear regressions, as well as the comparisons of the outputs for the general linear mixed models (GLMM), it is apparent that site, season, and year are strong influences on both of the independent variables I measured, which includes maximum height of an ecologically important marsh species, Pickleweed (*Salicornia pacifica*) and the average sediment deposition rates across a marsh bay ward edge. This is demonstrated from the six GLMMs performed, where the AIC values fell all within a close range, but only the model with all three random effects of Site, Season, and Year had the lowest value and was therefore the best fit model.

Based on the repeated outputs of coefficients for both distance and maximum height of Pickleweed, it is observed that both of these predictors are significant in terms of their effect on sediment deposition, but in very different ways.

There appears to be a negative relationship with distance and sediment deposition-meaning that the farther you go away from the marsh edge, the less sediment is deposited on the marsh platform. This makes sense from the ecological understanding, as the marsh edge is the commonly inundated with water than other points in the marsh, and the edge is more likely to encounter dynamic wave energy and scarps that suspended sediment in the water column would first interact with prior to settling out.

There also appears to be a positive relationship with maximum height and sediment deposition. This also makes sense ecologically, as a taller plant would create more of an obstruction to suspended sediment in the water column, causing for it to settle out quicker than with a shorter plant.

Both distance and maximum height are interrelated to other variables of the marsh, such as slope and average tide levels throughout the marsh. As demonstrated from the GLMM outputs, site specific characteristics such as slope and tide level cannot be ignored as predictors for sediment deposition rates, but they were not measured within this study. Slope is not accounted for in this model, and future analyses could take into account this additional variable as it could characterize the landscape differences between marshes that would result in such differences in both plant growth and sediment deposition. Slope is related to the mean low low water tide levels of the marsh, which are also pertinent variables to consider when measuring differences in sediment deposition as well as plant growth differences.

These results demonstrate that I can reject my null hypothesis based on the significance of the effect of distance and maximum height on differences in sediment deposition rates!

I am excited to see in my thesis studies how many of these additional variables I can measure to see if I can uncover at least one clear pattern with my own data-that would be incredible. Thanks for this class Jenna! I will sincerely miss it!
