---
title: "usgs_thesis_merge (Claude)"
format: html
editor: visual
---

Prep half of the old Comparing_USGS_to_Thesis.qmd, per the reorg plan -- this file only builds usgs_thesis.rds. Modeling (Shapiro-Wilk, GLMMs, ramped/scarped comparisons, etc.) moves to the repurposed "Research question 2.qmd", not duplicated here.

ADDED (this session): a replicate-level parallel throughout, producing usgs_thesis_byrep.rds/.csv alongside the existing averaged usgs_thesis.rds/.csv. Built from master_tier5_byrep.csv and usgs_bay_byrep.csv (the unaveraged outputs of step6_master_analysis_flagtier.qmd and usgs_step3_master_analysis_usgs_bayriver_FIXED.Rmd respectively), using the exact same cleanup/Month/Study.Dataset/merge/Flux-filter steps as the averaged version, just applied to the unaveraged pair. Neither the join logic nor the averaged output changed. Research_question_2_FIXED.qmd still needs to be pointed at usgs_thesis_byrep.rds and have (1\|Loc_ID) added to its models before it actually uses this -- that hasn't been done yet.

```{r, load packages}
source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\setup.R")
# No spatial_setup.R needed -- this file never calls sf/ggmap/leaflet,
# and it renders to html, not pdf, so tinytex isn't needed either.
```

```{r, load tier5 and usgs_bay from disk}
# Loads real data into properly-named objects -- the original file read
# into master_tier3/usgs_bay but then used sed_thesis/sed_usgs everywhere
# after, which were never created; it only "worked" by silently relying
# on objects left over in memory from another file's session. Fixed here
# by loading directly into the names the rest of this file actually uses.
#
# master_tier3.csv -> master_tier5.csv: under the current tier numbering,
# tier5 = (all sites, full flag criteria), which is what the old tier3
# meant before the tier renumbering earlier in this project.
#
# readr::read_csv() (not base read.csv()) so Date.placed.in.field /
# Date.removed.from.field come back as proper Date columns automatically,
# same reasoning as the read fix in master_analysis_cleanup.Rmd.

sed_thesis <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\master_tier5.csv")

sed_usgs <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_bay.csv")

str(sed_thesis)
str(sed_usgs)

# ADDED: replicate-level parallel, per the "don't average replicates"
# decision. master_tier5_byrep.csv / usgs_bay_byrep.csv are built by
# keep_replicates() in step6_master_analysis_flagtier.qmd and
# usgs_step3_master_analysis_usgs_bayriver_FIXED.Rmd respectively --
# same source pipelines as the averaged versions above, just the
# unaveraged output. Both already carry a Loc_ID column
# (interaction(Site, Tran, Dist)) built the same way on both sides, so
# it survives the merge below as a common column with no collisions --
# thesis Tran values ("1"/"2") and USGS bay Tran values ("A"/"B") never
# overlap, so Loc_ID stays unique across both datasets without needing
# Study.Dataset folded in.

sed_thesis_byrep <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\master_tier5_byrep.csv")

sed_usgs_byrep <- readr::read_csv("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_bay_byrep.csv")

str(sed_thesis_byrep)
str(sed_usgs_byrep)
```

```{r, clean whitespace and re-factor columns lost on CSV round-trip}
# CSVs don't preserve R factor levels, so both sides need re-factoring
# after the disk read -- this is real, necessary work, not duplication.
#
# NOT re-derived here (already present in both files from their upstream
# prep -- master_tier5.csv inherits marsh_edge_type/marsh_orientation
# from master_thesis via elev_z_flux; usgs_bay.csv already has
# marsh_edge_type/marsh_orientation/Study.Phase from master_analysis_usgs.Rmd):
#   - marsh_edge_type, marsh_orientation
#   - Study.Phase
# Re-deriving these again here would just be the same ~135 lines of
# case_when() logic duplicated a third time.

sed_thesis$Dist        <- str_replace_all(as.character(sed_thesis$Dist), " ", "")
sed_thesis$Tran         <- str_replace_all(as.character(sed_thesis$Tran), " ", "")
sed_thesis$Site         <- str_replace_all(sed_thesis$Site, " ", "")
sed_thesis$marsh_edge_type  <- as.factor(sed_thesis$marsh_edge_type)
sed_thesis$marsh_orientation <- as.factor(sed_thesis$marsh_orientation)
sed_thesis$Study.Phase  <- factor(sed_thesis$Study.Phase, levels = paste0("M", 1:12))

sed_usgs$Dist        <- str_replace_all(as.character(sed_usgs$Dist), " ", "")
sed_usgs$Tran         <- str_replace_all(as.character(sed_usgs$Tran), " ", "")
sed_usgs$Site         <- str_replace_all(sed_usgs$Site, " ", "")
sed_usgs$marsh_edge_type  <- as.factor(sed_usgs$marsh_edge_type)
sed_usgs$marsh_orientation <- as.factor(sed_usgs$marsh_orientation)
sed_usgs$Study.Phase  <- factor(sed_usgs$Study.Phase, levels = paste0("M", 1:12))

# Quick QC check before binding -- catches a site-naming mismatch here
# rather than discovering it as a silent NA later.
unique(sed_thesis$Site)
unique(sed_usgs$Site)

# ADDED: same re-factoring for the byrep pair.
sed_thesis_byrep$Dist        <- str_replace_all(as.character(sed_thesis_byrep$Dist), " ", "")
sed_thesis_byrep$Tran         <- str_replace_all(as.character(sed_thesis_byrep$Tran), " ", "")
sed_thesis_byrep$Site         <- str_replace_all(sed_thesis_byrep$Site, " ", "")
sed_thesis_byrep$marsh_edge_type  <- as.factor(sed_thesis_byrep$marsh_edge_type)
sed_thesis_byrep$marsh_orientation <- as.factor(sed_thesis_byrep$marsh_orientation)
sed_thesis_byrep$Study.Phase  <- factor(sed_thesis_byrep$Study.Phase, levels = paste0("M", 1:12))

sed_usgs_byrep$Dist        <- str_replace_all(as.character(sed_usgs_byrep$Dist), " ", "")
sed_usgs_byrep$Tran         <- str_replace_all(as.character(sed_usgs_byrep$Tran), " ", "")
sed_usgs_byrep$Site         <- str_replace_all(sed_usgs_byrep$Site, " ", "")
sed_usgs_byrep$marsh_edge_type  <- as.factor(sed_usgs_byrep$marsh_edge_type)
sed_usgs_byrep$marsh_orientation <- as.factor(sed_usgs_byrep$marsh_orientation)
sed_usgs_byrep$Study.Phase  <- factor(sed_usgs_byrep$Study.Phase, levels = paste0("M", 1:12))

unique(sed_thesis_byrep$Site)
unique(sed_usgs_byrep$Site)
```

```{r, derive Month for comparability}
# Derived from Date.removed.from.field rather than assumed to already
# exist as a raw column -- gives USGS/thesis comparability at the Month
# level, per your call to use Month (not the old Winter/Summer Season
# split) for this comparison.
#
# Factor levels ordered chronologically starting from July (the first
# month of data collection), not alphabetically -- so plots/models
# using Month read left-to-right in actual study order instead of
# April/August/December/... etc.

study_month_levels <- c("July", "August", "September", "October",
                         "November", "December", "January", "February",
                         "March", "April", "May", "June")

sed_thesis <- sed_thesis %>%
  mutate(Month = factor(month(Date.removed.from.field, label = TRUE, abbr = FALSE),
                         levels = study_month_levels, ordered = TRUE))

sed_usgs <- sed_usgs %>%
  mutate(Month = factor(month(Date.removed.from.field, label = TRUE, abbr = FALSE),
                         levels = study_month_levels, ordered = TRUE))

# ADDED: same Month re-derivation for the byrep pair.
sed_thesis_byrep <- sed_thesis_byrep %>%
  mutate(Month = factor(month(Date.removed.from.field, label = TRUE, abbr = FALSE),
                         levels = study_month_levels, ordered = TRUE))

sed_usgs_byrep <- sed_usgs_byrep %>%
  mutate(Month = factor(month(Date.removed.from.field, label = TRUE, abbr = FALSE),
                         levels = study_month_levels, ordered = TRUE))
```

```{r, Study Dataset column}
sed_thesis$Study.Dataset <- "Thesis"
sed_usgs$Study.Dataset   <- "USGS"

# ADDED
sed_thesis_byrep$Study.Dataset <- "Thesis"
sed_usgs_byrep$Study.Dataset   <- "USGS"
```

```{r, merge}
usgs_thesis <- bind_rows(sed_thesis, sed_usgs)

# bind_rows() fills any column that only exists on one side (e.g.
# z_star/Elevation_m, which only master_tier5 carries) with NA on the
# other side automatically -- expected, not a bug.

nrow(sed_thesis)
nrow(sed_usgs)
nrow(usgs_thesis)

# Safety filter -- both sides should already only have valid M1-M12
# Study.Phase values (usgs_bay.csv already dropped NA Study.Phase rows
# in master_analysis_usgs.Rmd), but keep this as a guard.
usgs_thesis <- usgs_thesis %>%
  filter(Study.Phase %in% paste0("M", 1:12))

unique(usgs_thesis$Study.Dataset)
unique(usgs_thesis$Study.Phase)
table(usgs_thesis$Study.Dataset, usgs_thesis$Month)

usgs_thesis %>% count(Study.Dataset, n_reps)

# ADDED: same merge + safety filter for the byrep pair.
usgs_thesis_byrep <- bind_rows(sed_thesis_byrep, sed_usgs_byrep)

nrow(sed_thesis_byrep)
nrow(sed_usgs_byrep)
nrow(usgs_thesis_byrep)

usgs_thesis_byrep <- usgs_thesis_byrep %>%
  filter(Study.Phase %in% paste0("M", 1:12))

unique(usgs_thesis_byrep$Study.Dataset)
unique(usgs_thesis_byrep$Study.Phase)
table(usgs_thesis_byrep$Study.Dataset, usgs_thesis_byrep$Month)

usgs_thesis_byrep %>% count(Study.Dataset, n_reps)
```

```{r}


# Applied once here, on the final merged dataset, so every downstream
# file (Research_question_2_FIXED.qmd, RQ3-variables.qmd, and any
# bay_channel/thesis_rq2 objects built from usgs_thesis) inherits the
# same filtered data automatically -- rather than needing filter(Flux >
# 0) re-applied inconsistently in each individual file, which is how the
# bay_channel Gamma GLMM error came up in the first place (bay_channel
# was built from usgs_bay.csv/usgs_river.csv directly, bypassing this
# step, and had 4 unfiltered zeros).
#
# Reported counts, not silently dropped -- worth knowing how many rows
# this removes and from where, for the methods write-up.

sum(usgs_thesis$Flux <= 0, na.rm = TRUE)
sum(usgs_thesis$Flux == 0, na.rm = TRUE)
sum(usgs_thesis$Flux < 0, na.rm = TRUE)

usgs_thesis %>%
  filter(Flux <= 0) %>%
  count(Study.Dataset, Site, Month)

# Save the full rows (every column, not just Flux) for the zero/negative
# entries BEFORE they're dropped -- gives you a standalone reference
# sheet to isolate exactly which trap/site/month/replicate combinations
# produced a zero, without having to re-run the merge or dig back
# through sed_thesis/sed_usgs to find them again later.

zero_flux_rows <- usgs_thesis %>%
  filter(Flux <= 0) %>%
  arrange(Study.Dataset, Site, Month)

nrow(zero_flux_rows)   # sanity check -- should match sum(usgs_thesis$Flux <= 0) above
zero_flux_rows

readr::write_csv(
  zero_flux_rows,
  "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis_zero_flux_rows.csv"
)

nrow(usgs_thesis)
usgs_thesis <- usgs_thesis %>%
  filter(Flux > 0)
nrow(usgs_thesis)
```

```{r, byrep zero flux check}
# ADDED: same Flux <= 0 diagnostic and filter for the byrep pair.
# Separate reference CSV (not overwriting the averaged version's) since
# the byrep grain will very likely flag more raw zero/negative rows --
# averaging can wash out a single bad replicate that keeping every
# trap separately won't.

sum(usgs_thesis_byrep$Flux <= 0, na.rm = TRUE)
sum(usgs_thesis_byrep$Flux == 0, na.rm = TRUE)
sum(usgs_thesis_byrep$Flux < 0, na.rm = TRUE)

usgs_thesis_byrep %>%
  filter(Flux <= 0) %>%
  count(Study.Dataset, Site, Month)

zero_flux_rows_byrep <- usgs_thesis_byrep %>%
  filter(Flux <= 0) %>%
  arrange(Study.Dataset, Site, Month)

nrow(zero_flux_rows_byrep)
zero_flux_rows_byrep

readr::write_csv(
  zero_flux_rows_byrep,
  "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis_byrep_zero_flux_rows.csv"
)

nrow(usgs_thesis_byrep)
usgs_thesis_byrep <- usgs_thesis_byrep %>%
  filter(Flux > 0)
nrow(usgs_thesis_byrep)
```

```{r, QC check against marsh area match guide}
# Non-blocking: warns instead of stopping if the guide is missing or a
# mismatch is found. This only validates usgs_thesis -- it never alters
# it or blocks the save below, even if this whole chunk fails.

crosswalk_check <- tryCatch({
  source("C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\code\\R\\marsh_area_crosswalk.R")
  build_marsh_area_crosswalk()
}, error = function(e) {
  warning("Marsh_Area crosswalk check skipped -- could not load guide: ", conditionMessage(e))
  NULL
})

if (!is.null(crosswalk_check)) {
  mismatch_check <- usgs_thesis %>%
    dplyr::left_join(
      crosswalk_check$long %>% dplyr::mutate(Dist = as.character(Dist)),
      by = c("Site", "Study.Dataset", "Dist"),
      suffix = c("", "_guide")
    ) %>%
    dplyr::filter(!is.na(Marsh_Area_guide),
                  as.character(Marsh_Area) != as.character(Marsh_Area_guide))

  if (nrow(mismatch_check) > 0) {
    warning(nrow(mismatch_check), " row(s) have a Marsh_Area that disagrees ",
            "with the match guide -- see mismatch_check for details.")
  } else {
    message("Marsh_Area crosswalk check passed -- no disagreements found with the match guide.")
  }

  # ADDED: same check against the byrep merge -- reuses crosswalk_check,
  # no need to rebuild the guide a second time.
  mismatch_check_byrep <- usgs_thesis_byrep %>%
    dplyr::left_join(
      crosswalk_check$long %>% dplyr::mutate(Dist = as.character(Dist)),
      by = c("Site", "Study.Dataset", "Dist"),
      suffix = c("", "_guide")
    ) %>%
    dplyr::filter(!is.na(Marsh_Area_guide),
                  as.character(Marsh_Area) != as.character(Marsh_Area_guide))

  if (nrow(mismatch_check_byrep) > 0) {
    warning(nrow(mismatch_check_byrep), " row(s) in usgs_thesis_byrep have a Marsh_Area that ",
            "disagrees with the match guide -- see mismatch_check_byrep for details.")
  } else {
    message("Marsh_Area crosswalk check passed for usgs_thesis_byrep -- no disagreements found.")
  }
}
```

```{r, save merged dataset}
saveRDS(usgs_thesis, "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis.rds")

readr::write_csv(usgs_thesis, "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis.csv")

# ADDED: replicate-level parallel, for Research_question_2_FIXED.qmd to
# switch to once it's updated to read this instead of usgs_thesis.rds.
saveRDS(usgs_thesis_byrep, "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis_byrep.rds")

readr::write_csv(usgs_thesis_byrep, "C:\\Users\\savan\\OneDrive - San Francisco State University\\ProjectData\\GitHub\\DepositionSFB\\data\\usgs_thesis_byrep.csv")
```
